---
title: "best_of_twitter"
---

# insert variables embedded with custom tags
https://twitter.com/moh_fodil/status/1684467674728779776?s=20


```{r}

var <- "john"
var_age <- 25

glue::glue(
  '{
  "name":"<<<var>>>",
  "age":"<<<var_age>>>"
  
  }',
  .open = "<<<"
  ,.close=">>>"
)


```

## case when with tibbles

https://twitter.com/antoine_fabri/status/1683557715115687939?s=20

Can add multiple columns / tibbles without a named argument

```{r}
cars %>% 
  mutate(case_when(
    speed==20~ tibble(a=1,b=2)
    ,speed==7 ~ tibble(a=3)
  )
  )

```


R and Github action
https://www.youtube.com/watch?v=u-TcADhWosA&ab_channel=RfortheRestofUs    

http://haines-lab.com/post/2022-01-23-automating-computational-reproducibility-with-r-using-renv-docker-and-github-actions/

#functional programming


# data dictionary

https://tim-tiefenbach.de/post/2023-recode-columns/

# intro to R
https://www.youtube.com/playlist?list=PLHfe4A2DV-b2XDFl-9ytrcmQVZ9NNOEQS


https://www.marvinschmitt.com/blog/functional-programming-r-basics/?card=true

## .Last.value

-   whatever last analysis you run you can re-run it again with `.Last.value`

-   helpful when you have done complex analysis and don't want to waste time re-running it

```{r}
x <- 1:10


.Last.value
```

