---
title: "database"
---


## Resource
https://smithjd.github.io/sql-pet/chapter-setup-adventureworks-db.html#
https://r4ds.hadley.nz/databases.html


# How to connect

-   First need to configure your DSN so that you can be authenticated
-   If done correctly you can then use DBI package to connect to the database of choice

        -   create a connection string with DBI::dbConnect()
            -   driver_name such as ODBC::ODBC()
            -   DSN name
        -   Enter SQL query with `sql()`
        -   Use connection string and sql query together to create a lazy table with `tbl()`
-   From there you can use dplyr back end queries to see everything
-   You can see what query it will generate with `show_query()`
-   You can finallly execute the query with `collect()`

-   if you already have something as a dataframe you can upload it to a database with DBI::dbWriteTable(con,"tbl_name",df) which will write the table to the connection with the name you gave

    -   dbWriteTable() can write a r dataframe or you can use sql to create a virtual table if you want
    
    

-   dbListTables(con) can be used to see what virtual tables are in the connection

- dbReadTable(con,"tbl_name") read the virtual table as a dataframe

- in_schema("database","tbl") is a way to navigate a scehma in tbl(

con <- DBI::dbConnect(ODBC::ODBC(),dsn="name")

```{r}

```



# special example to pass one db object to another

```{r}
#| echo:fenced
#| label: db-connect example
#| eval: false

library(DBI)
library(tidyverse)

## create connection string locally
con_db <- DBI::dbConnect(duckdb::duckdb())

# loads data into your connection either in memory
dbWriteTable(con_db,"diamonds_db",ggplot2::diamonds)

#create new table to the connection

dbExecute(con_db, "CREATE TABLE duckdb_table (col1 INT, col2 STRING)")

#preview what is in your connection

dbListTables(con_db)



diamonds_db <- tbl(con_db
    ,"diamonds_db"
    )


tbl(con_db
    ,"duckdb_table"
    )

```






```{r}
diamonds_db %>% 
  
  mutate(good_indicator=if_else(cut=="Good",1,0)) %>%
  group_by(color) %>% 
  summarise(
    n=n()
    ,mean=mean(carat)
    ,mean_price=mean(price)
    ,mean_ind=mean(good_indicator)
    ,mean_adj=mean(carat[good_indicator==1])
  ) %>% 

  arrange(desc(color)) %>% 
  mutate(rolling_price=cumsum(mean_price)) %>% 
    arrow::to_arrow() %>% 
  ungroup() %>% 
 filter(color=="H") %>% 
  select(color,mean_adj) 
  arrow::to_duckdb()
```

```{r}
diamonds_db %>%
  collect() %>% 
  nest_by(cut) %>% 
  mutate(
    data=list(data)
    ,mod=list(lm(price ~ carat+x,data=data))
    ,gl_mod=list(broom::glance(mod))
    ) %>% 
  unnest(gl_mod)

library(sparklyr)

sc <- spark_connect(master="local")
test <-  sparklyr::copy_to(sc,diamonds,"diamonds_spk",overwrite = TRUE)



spark_apply(
  test
   ,nrow
    ,group_by="cut"
    )

```

```{r}
library(tidyverse)
install.packages("skimr")

diamonds %>% 
  names()
diamonds %>% 
  skimr::skim() %>% 
  write_csv(.,"test.csv")


labelled::generate_dictionary
)
remotes::install_github("REditorSupport/languageserver")
languageserversetup::languageserver_startup()
languageserversetup::languageserver_install()
yesremotes::install_github("REditorSupport/languageserver")
languageserversetup::languageserver_add_to_rprofile()
install.packages("styler")
yeslanguageserversetup::languageserver_remove_from_rprofile()
sessionInfo()
install.packages("languageserver")
library(styler)
library(languageserversetup)
languageserver::run()


```

```{r}

df <- structure(list(Pitcnt = c("12 (3-2)", "7 (0-1)", "3 (0-2)", 
"8 (3-2)", "5 (2-2)", "7 (3-2)", "1 (0-0)", "2 (0-1)", "5 (1-2)", 
"10 (3-2)")), class = c("tbl_df", "tbl", "data.frame"), row.names = c(NA, 
-10L))

library(tidyverse)

df %>% 
  separate(col = Pitcnt, into=c("1","2","3"))

```

