---
title: "data_masking"
---

## Resources

[https://brad-cannell.github.io/r_notes/confidence-intervals.html]


### data masking


enquoes
-   quotes functions eg. turns mpg into "mpg" AND declares the environment eg. global or local for the quoted return to be put in. 
-   eg. ... or mpg needs to but into enquos(...) or enquo(mpg)
-   you evalute unquoted objects with !! or despare
- or combine both enquo() and !! with {}
-   if you have a list of quoted objects you need !!! to unlist

- rlang::qq_show() around your evaluated expresssion to show
-   if input is a quoted input put it around sym (enquo for unquoted)



```{r}
library(tidyverse)

my_mean <- function(data, var) {
  
  var <- ensym(var)
  
  rlang::qq_show(data %>% dplyr::summarise(mean(var)))
}

mtcars %>% my_mean(cyl)
```


```{r}

custom <- function(.data,group,focus){
  
  group <- enquos(group)
  focus <- enquo(focus)
    

  
    .data %>% 
    group_by(pick(!!!group)) %>% 
      summarize(
        mean=mean(!!focus)
        ,.groups="drop"
      )
  
  
}

diamonds %>% custom(group=c(color,cut),focus=x)
```



sym() is used to take a quoted input "x" and use it in a context without quotes. Need !!sym("x") to be used

or  rlang::englue("var) within the .data[[var]] context


.data[[rlang::englue("var")]]

.env$var to refer to grid table

```{r}
library(tidyverse)


col_numbers <- 1:ncol(diamonds)

  map(
    .x=col_numbers
      ,.f=~diamonds[.x:10]
  )

  split(diamonds, ~ cut) %>% 
    map(
      .x=.
      ,~.x %>% pull(price) %>% sum) %>% 
    stack()

 
    map(
      .x=col_numbers
      ,.f = ~mtcars[.x] %>% table()
      )
```



```{r}
diamonds %>% select(cut,color) %>% 
  table()
```



```{r}
library(tidyverse)
cols <- c("mpg","cyl")

df <- expand.grid(cols) %>% 
  mutate(data=list(mtcars))


df
global_env_unquoted <- function(data,col){
  
  
  data %>% 
    filter(eval({{col}}))
}



df %>% 
  rowwise() %>% 
  mutate(args=list(rlang::parse_expr(paste0(Var1,">18")))) %>% 
  mutate(
    fun=list(global_env_unquoted(data,args))
  ) 

```


