---
title: "database"
---


## Resource
https://smithjd.github.io/sql-pet/chapter-setup-adventureworks-db.html#
https://r4ds.hadley.nz/databases.html


# How to connect

-   First need to configure your DSN so that you can be authenticated
-   If done correctly you can then use DBI package to connect to the database of choice

        -   create a connection string with DBI::dbConnect()
            -   driver_name such as ODBC::ODBC() or duckdb::duckdb()
            -   DSN name (if applicable)
        -   Enter SQL query with `sql()`
        -   Use connection string and sql query together to create a lazy table with `tbl()`
-   From there you can use dplyr back end queries to see everything
-   You are limited to queries that can be translated to 
-   You can see what query it will generate with `show_query()`
-   You can finallly execute the query with `collect()`

-   if you already have something as a dataframe you can upload it to a database with DBI::dbWriteTable(con,"tbl_name",df) which will write the table to the connection with the name you gave

    -   dbWriteTable() can write a r dataframe or you can use sql to create a virtual table if you want
    -   dbplyr::copy_inline(con_db,df = df) is alternative method
    
-   dbListTables(con) can be used to see what virtual tables are in the connection

- dbReadTable(con,"tbl_name") read the virtual table as a dataframe

- in_schema("database","tbl") is a way to navigate a scehma in tbl()

con <- DBI::dbConnect(ODBC::ODBC(),dsn="name")

con_db <- DBI::dbConnect(duckdb::duckdb())






# special example to pass one db object to another

```{r}
#| echo: fenced
#| label: db-connect example
#| eval: true

library(tidyverse)

## create connection string locally
con_db <- DBI::dbConnect(duckdb::duckdb())

# loads data into your connection either in memory
DBI::dbWriteTable(con_db,"diamonds_db",ggplot2::diamonds)

#create new table to the connection

DBI::dbExecute(con_db, "CREATE TABLE duckdb_table (col1 INT, col2 STRING)")

#preview what is in your connection

DBI::dbListTables(con_db)

dbplyr::copy_inline(con_db,df = diamonds)


diamonds_db <- dplyr::tbl(con_db
    ,"diamonds_db"
    )


print(diamonds_db)

```



# Example of passing one database object to arrow


```{r}
#| echo: true
#| label: query-example
#| eval: true
#| error: false
#| warning: false


diamonds_db %>% 
  
  mutate(good_indicator=if_else(cut=="Good",1,0)) %>%
  group_by(color) %>% 
  summarise(
    n=n()
    ,mean=mean(carat)
    ,mean_price=mean(price)
    ,mean_ind=mean(good_indicator)
    ,mean_adj=mean(carat[good_indicator==1])
  ) %>% 

  arrange(desc(color)) %>% 
  mutate(rolling_price=cumsum(mean_price)) %>% 
    arrow::to_arrow() %>% 
  ungroup() %>% 
 filter(color=="H") %>% 
  select(color,mean_adj) %>% 
  collect()
```



