---
title: "database"
---

## Resource
https://smithjd.github.io/sql-pet/chapter-setup-adventureworks-db.html#
https://r4ds.hadley.nz/databases.html
https://dbplyr.tidyverse.org/articles/translation-function.html
https://bwlewis.github.io/duckdb_and_r/talk/talk.html


# What do you need?
-   Probably the most frustrating part of database is getting access to the database and making the initial connection
-   There are probably multiple ways to connect almost all require the following
    -   user name
    -   password
    -   database driver
    -   connection string and associated arguments
    -   SQL query

# How to connect (DSN method)

## Set up ODBC Driver
-   First need to configure your DSN so that you can be authenticated
-   Your database platform should have documentation on how to do this and your internal IT team should be able to articulare any proxy / security requirements
-   Here is an example of [documentation](https://docs.snowflake.com/en/developer-guide/odbc/odbc-windows)

## Create Connection String
-   If done correctly you can then use DBI package in R to connect to the database of your choice

        -   create a connection string with DBI::dbConnect()
            -   Select the DBMS wtih the driver_name such as ODBC::ODBC() to access your DSN set up connect to external database or can use DMBS package such as duckdb::duckdb() to replicate an internal instance of a database
            -   DSN name if external database (the name used to set up ODBC driver)
            -   Alternatively, you can directly supply the arguments in DBI::dbConnect() such as hostname,port,username, etc

```{r}
#| echo: true
#| label: name
#| eval: true
#| error: false
#| warning: false


con <- DBI::dbConnect(duckdb::duckdb())


```

## Option 1: Create SQL string

-   If you know the database, schema and table name that you want, you can direct write an initial sql query to connect to the database
        
        -   Enter SQL query with `dplyr::sql()`
        
```{r}
#| echo: true
#| label: name
#| eval: false
#| error: false
#| warning: false


sql_script <- "select *  from database_name"
```

## Accessing Databse

-   Use connection string and sql query together to create a lazy table with `dplyr::tbl()`

```{r}
#| echo: true
#| label: name
#| eval: false
#| error: false
#| warning: false

data_db <- tbl(con,sql_script)
```

-   From there you can use dplyr back end queries to see everything
-   You are limited to queries that can be translated to sql which are detailed below
    -   [github of dplyr commands that can be used in dbplyr](https://github.com/tidyverse/dbplyr/blob/main/R/backend-.R)
    -   You can also check the database specific[here]([https://github.com/tidyverse/dbplyr/blob/main/R/backend-snowflake.R)
-   You can see what query it will generate with `show_query()`
-   You can finally execute the query with `collect()`


## Existing data into database
-   if you already have something as a dataframe you can upload it to a database with DBI::dbWriteTable(con,"tbl_name",df) which will write the table to the connection with the name you gave 

    -   dbWriteTable() can write a r dataframe or you can use sql to create a virtual table if you want
    -   dbplyr::copy_inline(con_db,df = df) is alternative method

```{r}
library(tidyverse)

# create connection
con_db <- DBI::dbConnect(duckdb::duckdb())

# write data into database
DBI::dbWriteTable(con_db,name = "diamonds_db",value = ggplot2::diamonds)
# or alternative use the database argument to regester
duckdb::duckdb_register(con_db, "diamonds_db_2",df =  ggplot2::diamonds)
# validate data is in database by reference connection
DBI::dbListTables(con_db)

# Pull in data in database format
diamonds_db <- dplyr::tbl(con_db,"diamonds_db")

# 

```

## Larger than memory problems


-   Somtimes you may not have a database to connect to and instead have a very large csv file
-   If you have large offline files you can quickly easily load this into duckdb with `duckdb_read_csv()` very helpful to manipulate larger than memory problems
-   similar trick exists with arrow which converts it to a parquet file

# special example to pass one db object to another

```{r}
#| echo: fenced
#| label: db-connect example
#| eval: true

library(tidyverse)

## create connection string locally
con_db <- DBI::dbConnect(duckdb::duckdb())

# loads data into your connection either in memory
DBI::dbWriteTable(con_db,"diamonds_db",ggplot2::diamonds)

#create new table to the connection

DBI::dbExecute(con_db, "CREATE TABLE duckdb_table (col1 INT, col2 STRING)")

#preview what is in your connection

DBI::dbListTables(con_db)

dbplyr::copy_inline(con_db,df = diamonds)


diamonds_db <- dplyr::tbl(con_db
    ,"diamonds_db"
    )


print(diamonds_db)


```



# Example of passing one database object to arrow


```{r}
#| echo: true
#| label: query-example
#| eval: true
#| error: false
#| warning: false


diamonds_db %>% 
  
  mutate(good_indicator=if_else(cut=="Good",1,0)) %>%
  group_by(color) %>% 
  summarise(
    n=n()
    ,mean=mean(carat)
    ,mean_price=mean(price)
    ,mean_ind=mean(good_indicator)
    ,mean_adj=mean(carat[good_indicator==1])
  ) %>% 

  arrange(desc(color)) %>% 
  mutate(rolling_price=cumsum(mean_price)) %>% 
    arrow::to_arrow() %>% 
  ungroup() %>% 
 filter(color=="H") %>% 
  select(color,mean_adj) %>% 
  collect()
```



## how to view schema / catalog of a database
tbl(con,in_schame(schema_name,database_name))
## Useful tricks to  dynamically render a sql

## automating databse queries / reporting 

    -     Scheduler
    -     Cron tab
    
    
## Show query output results 

## DBI connection useful utilities

https://dbi.r-dbi.org/
https://solutions.posit.co/connections/db/

### list tables in connection
dbListTables to list tables associated with a connection

### list drivers
sort(unique(odbcListDrivers()[[1]]))

### Running SQL in R

```\{sql, connection=con, output.var = "mydataframe"\}
```

## Run SQL queries safely


## visualizations

dbplot_histogram

(https://edgararuiz.github.io/dbplot/)


