---
title: "Database and Larger Than Memory Problems"
html:
  code-fold: "show"
---



# Resource

[Resource to understand databases](https://smithjd.github.io/sql-pet/chapter-setup-adventureworks-db.html#)

[R for datascience chapter on database](https://r4ds.hadley.nz/databases.html)

[Dbplyr functions](https://dbplyr.tidyverse.org/articles/translation-function.html)

[How to use duckdb](https://bwlewis.github.io/duckdb_and_r/talk/talk.html)

[DBI function overivew](https://dbi.r-dbi.org/)

[Posit learning materials](https://solutions.posit.co/connections/db/)

[dbplot_histogram](https://edgararuiz.github.io/dbplot/)



# Why is this important?

-   You will get to a point where the data you need is inside a database and 
not excel sheets or csv files 

-   This is a point of potential rejoice or mourning

-   While we can say goodbye excel sheets and csv files we also need to say hello 
to poorly documented databases, overwhelmed database product owners,
overworked data engineers, and finally SQL and all of its variations

-   I recommend you take the time to learn SQL -- the basics are very
similar to the dplyr commands you already know (slight tweaking of 
evaluation order and syntax) so the beginners learning curve isn't steep

:::{.callout-note collapse="true"}

## How much SQL should I learn?

If you're curious how much SQL you should learn, below is a framework that I
found to be helpful, with their Rstats comparison

    
```{r}
#| warning: false
#| echo: false

library(tidyverse)

tribble(
  ~sql,~dpylr,~comment
  ,"where"               ,"filter()", "can be viewed as one for one replacement to filter(), you can pass arguments with AND or OR,\nSometimes the first argument is TRUE for purely formatting purposes"
  ,"select"             ,"select()" , "Can be viewed as one for one replacement for select(), * is short cut for all columns"
  ,"group by"           ,"group_by()", "can be viewed as one for replacement for group_by(), you need for it summarization measures as well as partition measures "
  ,"sum"                ,"sum()","works the same as sum, na.rm is always TRUE for SQL"
  ,"min"                ,"min()","works the same as sum, na.rm is always TRUE for SQL" 
  ,"max"                ,"max()","works the same as sum, na.rm is always TRUE for SQL"
  ,"count(*)"           ,"n()","works similiar to n()" 
  ,"select distinct"    , "select() |> distinct()","works almost exactly the same"
  ,"partition"          ,"group_by() |> mutate()","works exactly the same "
  ,"join"               , "left_join()", "works similiar"
  ,"like"               ,"str_detect()", "similiar"
  ,"top"                ,"head()"      ,"works the same"
  ,"DISTINCT"           ,"distinct()"  ,"works the same"
  ,"DESCRIBE"           ,"glimpse()"   ,"works simliar, will display the columns and thier data types"
  ,"SET"                ," <- "        ,"works simliar to assign single variables"
  ,"BETWEEN"            ,"between()"   ,"simliar"
  ,"MONTH"              ,"month()"     ,"simliar"
  ,"YEAR"               ,"year()"      ,"simliar"
  ,"DAY"                ,"day()"       ,"simliar"
  ,"QUARTER"            ,"quarter()"   ,"simliar"
  ,"DATEDIFF"           ,"difftime() "   ,"simliar"
  ,"DATETRUC"           ,"floor_date()","similar"
  ,"create or replace"  ,"tibble()","use this to create the final table"
  ,'in'                 ," %in%", "similiar"
  ,"as"                 , "rename()","simliar"
  ,"with"               , "tibble()", "Use this to create CTE or basic mini tables that you can then reference in different steps of the query, makes readable easier"
  ,"having"             , "filter() or '['"," similiar to where"
) |> 
  gt::gt() |> 
  gt::cols_label(
    sql="SQL"
    ) |> 
  gt::tab_header(
    title="Summary of SQL commands and their dplyr counterpart"
  )


```
    
There are some nuances, in particular the evaluation order that typically trip up new SQL users and also some coding conventions that will be different than what you are used to but in general if you understand the above R commands you will quickly learn the SQL counterparts.

With anything, you need to practice! luckily there are multiple SQL resources and practice studios which can help with reinforcement learning.


**Database frameworks**


-   Temporary Tables / CTE
    
    -   These are tables that only exist when you run them
    -   Helpful as interim steps or to break code into subqueries to make it more modular

-   Curated tables
    
    -   Often times you may have loads of raw tables (eg 100s) that you need to join together, filter or aggregrate before the data can be usable
    -   This process of turning raw /streaming data into table that can be consumed for analysis is oftern called data curation
    -   This is often times created as a view which can be though of as particular snapshot of a table 
-   Materialized layers
    
    -   Materialized layer means the data is more persistent so when you run it its not triggering the underlying queries (which will save you alot of time)      
    -   Typically you won't know 


Database structure

-   Security Model

    -   Because data can be privildedged, without a doubt your organization has some security model that will aplly row level security and IDs to ensure when you access a table you are seeing what you should be seeing
    -    There is way to  much to write here about it and honestly, I'm not the right person to answer it 
  


You may not need to know any of this but this mostly depends on your organizations
data strategy, staffing levels and operating model

:::


-   The reason you can survive off of a beginners knowledge base of SQL is that
luckily there is a life saving package called `dbplyr` that translates your 
dplyr queries into SQL for you

-   It has fairly great coverage but there are limitations which is why 
eventually it will help you to learn some intermediate SQL commands and 
overall database frameworks

-   This chapter will go over database essentials and provide resources to
learn more

## The Essentials

**What do you need to access data in a database?**

-   A database with data inside of it  
-   Access / permission to the database
-   Location, user name and password to database (or equivalent protocols as 
dictated by your organization's security model)
-   Database driver and related utilities
-   SQL querys
-   Patience


::: {.callout-note collapse="true"}

## "What is the advantage of a database?" 

It comes down to scale and size. At some point your organization or process
will generate substantial data and needs a more structured process to store
the data so that multiple parties can access the data at scale.

When dealing with a new database some key frameworks are:

- Cloud vs. On-Premise
- Security model and access
- "Flavor" of database


- Improved data management: A database centralizes data, making it 
easier to manage and maintain

- Enhanced data security: A database provides secure storage and 
retrieval of sensitive data through user authentication and access 
control mechanisms

- Better data organization: A database **theoretically** allows for better organization 
and structure of data through the use of tables, indexes, and 
relationships

- Improved query performance: Databases are optimized for query 
performance, allowing for faster retrieval of data.


- Scalability: Databases can handle large amounts of data and scale as 
needed to meet growing storage demands.

Cloud vs On-Premise Database:

- Cloud databases are hosted on remote servers, while on-premise 
databases are hosted on local servers (when you read servers just replace it with the word computers. You are either using your organization's computer (on prem) or you are using someone else (cloud))

- Cloud databases offer greater flexibility in terms of scalability and
accessibility, as they can be accessed from any location with an 
internet connection.

- On-premise databases provide more control over data security and 
privacy, as the data is stored on a local server and not transmitted 
over the internet.

- Cloud databases typically require less setup and maintenance than 
on-premise databases, as they are managed by the provider.

- Cost: Cloud databases are often subscription-based and can be more 
cost-effective than on-premise databases, especially for small to 
medium-sized businesses.

*Security Model and Access:*

- Security model: A database's security model determines who has access
to the data and how they can access it. Common security models include 
Role-Based Access Control (RBAC), Attribute-Based Access Control (ABAC),
and Identity-Based Access Control (IBAC).

- Access control: A database's access control mechanisms determine who 
can view, edit, or delete data. This can be based on user 
authentication, role-based access control, or attribute-based access 
control.


- Authentication methods: Databases support various authentication 
methods such as username and password, single sign-on (SSO), and 
two-factor authentication (2FA).

- Authorization methods: Databases support various authorization 
methods such as row-level security, column-level security, and 
table-level security.

- Auditing and logging: Databases can log all access attempts and 
successful accesses to track user activity and detect potential security
breaches.

- Encryption: Databases can encrypt data both in transit and at rest to
protect it from unauthorized access.

- Backup and recovery: Databases provide mechanisms for backing up data
and recovering from failures or security incidents.

- Identity and access management (IAM): IAM systems manage user 
identities and access rights within the database, ensuring that only 
authorized users can access the data.

- Role-based access control (RBAC): RBAC allows for assigning roles to 
users based on their job function or responsibilities, limiting the data
they have access to.

- Attribute-based access control (ABAC): ABAC grants or denies access 
to data based on attributes associated with the user or the data itself,
such as location or time of day.

:::


**What does a database need from you?**

-   The most frustrating part of database is getting access to the database,
setting up the database utilities and then making the initial connection

-   There are multiple ways to connect to a database however  almost all
require the following:
    -   User name
    -   Password
    -   Database driver
    -   Connection string and associated arguments
    -   SQL query

-   We will review the DSN method for connecting to a database

-   These tend to be confusing because much of this is controlled and managed
by your local IT department so whatever documentation or guide you read on 
online may not translate one for one to your localized experienced (this 
includes this guide as well :(

## Set up ODBC Driver

-   Download (if required) a database driver for your database -- this is
typically on the database company's website

    -    Your company may have centralized package manager system where you
    will need to download and install all required drivers via that packet 
    manager

-   Configure your DSN so that you can be authenticated

-   Your database platform should have documentation on how to do this and 
your internal IT team **should** be able to articulate any proxy / security 
requirements

-   Here is some example
[documentation](https://docs.snowflake.com/en/developer-guide/odbc/odbc-windows)

-   You are essentially saving the required information (listed above)
to your computer so that you can pass these arguments to the database
-   Pay attention to the name you setup the DSN driver, you will need this
later one

**Example Paramaters are below**

-   User
-   Password
-   Server
-   Database
-   Schema
-   Warehouse      
-   Tracing
-   Authenticator
  

## Create Connection String

-   If the above is done correctly you can then use DBI package in R 
to connect to the database of your choice

-  Create a connection string with DBI::dbConnect()

      -   Select the DBMS wtih the driver_name such as ODBC::ODBC() 
            to access your DSN set up connect to external database or can use
            DMBS package such as duckdb::duckdb() to replicate an internal 
            instance of a database

      -   DSN name if external database (the name used to set up ODBC driver)

      -   Alternatively, you can directly supply the arguments 
            in DBI::dbConnect() such as hostname,port,username, etc



```{r}
#| echo: true
#| eval: false
#| code-fold: true
#| error: false
#| warning: false
#| fig-cap: "hello"
#| label: "fig-example"

## this uses duckdb example to create a connection string

con <- DBI::dbConnect(drv=duckdb::duckdb())

## this is alternative example using a made up DSN name 

con  <- DBI::dbConnect(
  drv=odbc:odbc()
  ,dsn="your_DSN_name" 
  )

```


::: {.callout-note collapse="true"}

## Additional Utilities

-   the DBI and ODBC packages are extremely useful for database related 
utilities
-   While they have some existing overlap, they can be used to view the 
schema in your database, list active connections and also disconnect.
-   Below are some useful utilities:


#### list your drivers
odbc::odbcListDrivers()





#### list active connections

DBI::dbListConnections()


## Checks if you can connect

DBI::dbCanConnect()

## List tables listed under connection

DBI::dbListTables()

### list tables in connection
`dbListTables(con)` to list tables associated with a connection

:::

-   After you have created a connection string you now need to retrieve 
information from the database

   
## Option 1: Create SQL string

-   If you know the database, schema and table name that you want, you can 
write the initial sql query to connect to the database

-   You can write simple or advance query insde the  `dplyr::sql()` function 
        
```{r}
#| echo: true
#| eval: false
#| error: false
#| warning: false


sql_query <- dplyr::SQL("select *  from database_name")
```

## Accessing Databse

-   Use connection string and sql query together to create a lazy table with `dplyr::tbl()`
-   We call this a lazy table because it won't actually execute the query and 
return the results which is good because your query might return 100s of results

```{r}
#| echo: true
#| eval: false
#| error: false
#| warning: false

data_db <- dplyr::tbl(con,sql_query)
```

-   From there you can use *dplyr* back end queries to see everything 
(notice the distinction betwee **db**pyr and dplyr

-   If you use the dbplyr package, you are limited to queries that can 
be translated to sql which are detailed below
    -   [github of dplyr commands that can be used in dbplyr](https://github.com/tidyverse/dbplyr/blob/main/R/backend-.R)
    -   You can also check the database specific[here]([https://github.com/tidyverse/dbplyr/blob/main/R/backend-snowflake.R)
-   You can see what query it will generate with `dbplyr::show_query()`
-   Notice the class of the object you return, you are returning a database 
object -- if you want to return a dataframe you need to use  `dplyr::collect()`


## Option 2: Push existing data into a database

-   First you need to have a connection string to a database (and write permissions)
    
-   If you already have something as a dataframe you can upload it to a 
database with DBI::dbWriteTable(con,"tbl_name",df) which will write the table 
to the connection with the name you gave 

    -   DBI::dbWriteTable() can write a r dataframe or you can use sql to create a virtual table if you want
    -   dbplyr::copy_inline(con_db,df = df) is alternative method
    -   If you have duckdb connection you can use the duckdb::duckdb_register()
```{r}
# create connection
con_db <- DBI::dbConnect(duckdb::duckdb())

# write data into database
DBI::dbWriteTable(con_db,name = "diamonds_db",value = ggplot2::diamonds)

# or alternative use the database argument to regester
duckdb::duckdb_register(con_db, "diamonds_db_2",df =  ggplot2::diamonds)

# validate data is in database by reference connection
DBI::dbListTables(con_db)

# Pull in data in database format
diamonds_db <- dplyr::tbl(con_db,"diamonds_db")

```

## What happens if dbplyr doesn't have a function that I need?

### built in helper functions

### create sql query and use it


## how can I build a package for this?

### sprintf() and SQL

### dbplyr

## Putting it all together

-   Set up your drive, get required database info and related utilities
-   Create connection to your database
-   write an inital query to select the columns that you want or need
-   Use dbplyr to translate dplyr queries to SQL
-   return results to your local machine with dplyr::collect()

# Larger than memory problems

-   Sometimes you don't have a database but have data that is larger than
memory
-   Luckily, you don't need a database to take advantage of the tricks we have 
learnt to move solve your data larger than memomory problems


-   Sometimes you may not have a database to connect to and instead have a 
very large csv file or dataframes
-   There are two packages that are exceptionally helpful here
    -   duckdb() and arrow()
-   This isn't the technically correct response but duckdb() allows you to build
an in memory database whereas arrow compresses your information into some sort 
of parquet type structure

-   What makes these so great is that dbplyr will translate your dplyr commands
to either duckdb or arrow language
-   Some dplyr functions are avaialble in one package that aren't avaialble in 
the other
-   However you can pass an object from duckdb to arrow as much as you want

**What is key difference?**

-   Duckdb will return first 1000 rows of your query so you can check if your
query worked well, whereas arrow won't let you see it (including if your query
returns an error which can be annoying)


To understand how to use the packages,let us define two scenarios: 

1) You have a lot of csv files that you need to upload and analyze
2) As a result of simulation of some other you have multiple tables that 
seperately are okay but together are generating you have very large
dataframe that you need to join together to  analyze and manipulate
3) You have interim data that you want to save in workflow that is large

### Large csv files that you need to work with:

-   If you have large offline files you can quickly easily load this into 
duckdb with `duckdb_read_csv()` or the collection of arrow functions 
(eg. `arrow::read_csv_arrow()`).


::: .callout-note

Arrow will support json, feather, delimiter, parquet, or csv amongst other 
whereas duckdb only supports csv (if not already an object) 
:::

-   This will load the csv files directly to your duck db in memory database
-   You first need to create connection string

### Existing dataframes (however they got there) that are either too large
or individually are okay but seperately aren't okay
   
-   You first need to get your data into R as a dataframe 
-   Then you need to register your dataframe to your in duckdb memory database
-   From there you can move the datafrme from duckdb to arrow as you would like



# Special tricks

##  You can pass one database object to duckdb


```{r}
#| echo: fenced
#| label: db-connect example
#| eval: true


## create connection string locally
con_db <- DBI::dbConnect(duckdb::duckdb())

# loads data into your connection either in memory
DBI::dbWriteTable(con_db,"diamonds_db",ggplot2::diamonds)

#create new table to the connection

DBI::dbExecute(con_db, "CREATE TABLE duckdb_table (col1 INT, col2 STRING)")
```
#preview what is in your connection
```{r}
#| eval: false

DBI::dbListTables(con_db)

dbplyr::copy_inline(con_db,df = diamonds)


diamonds_db <- dplyr::tbl(con_db
    ,"diamonds_db"
    )



```


# Example of passing one database object to arrow

```{r}
#| echo: true
#| label: query-example
#| eval: false
#| error: false
#| warning: false


diamonds_db %>% 
  
  mutate(good_indicator=if_else(cut=="Good",1,0)) %>%
  group_by(color) %>% 
  summarise(
    n=n()
    ,mean=mean(carat)
    ,mean_price=mean(price)
    ,mean_ind=mean(good_indicator)
    ,mean_adj=mean(carat[good_indicator==1])
  ) %>% 

  arrange(desc(color)) %>% 
  mutate(rolling_price=cumsum(mean_price)) %>% 
    arrow::to_arrow() %>% 
  ungroup() %>% 
 filter(color=="H") %>% 
  select(color,mean_adj) %>% 
  collect()
```

# Additional tricks

### Running SQL in R

-   If you are using rmakrdown or quarto, you can run the sql query in a window
and have it results saved as a datafarme

```{sql}
--| eval: false
--| connection: con_db
--| output.var: test.tbl
--| echo: fenced

SELECT * 

FROM diamonds_db

WHERE cut=='Good'

LIMIT 100



```

-   If you want to run it in rmarkdown, you can do the following

```{sql, eval=FALSE,connection=con_db, output.var = "mydataframe"}
--| echo: fenced
SELECT * 

FROM diamonds_db

WHERE cut=='Good'

LIMIT 100

```

## Run SQL queries safely

## How to plot a database object


## How to build packages or function to access dbplyr


-   two options:
    
    -  sprintf to generate actual sql queries
    -   dbplyr / similiar to spring
