---
title: "functions"
---

```{r}
get_symbols <- function() {
  wheel <- c("DD", "7", "BBB", "BB", "B", "C", "0")
  sample(wheel, size = 3, replace = TRUE, 
    prob = c(0.03, 0.03, 0.06, 0.1, 0.25, 0.01, 0.52))
}


get_symbols()


payouts <- c("DD" = 100, "7" = 80, "BBB" = 40, "BB" = 25, 
  "B" = 10, "C" = 10, "0" = 0)

payouts




list <- map(
  .x = c("2023-1-01","2022-12-02","2022-01-30")
  ,.f= function(date) seq.Date(from=ymd("2022-01-01"),to = ymd(date),by = "days")
  ) 


list[[1]][[3]]


roll <- function(){
  
 dice <- 1:6
 roll <-  sample(dice,2,replace = TRUE)
 sum(roll)
  
}


roll()

```

## subsetting


How to do lookup tables in R

```{r}
score <- function (symbols) {
  # identify case
  same <- symbols[1] == symbols[2] && symbols[2] == symbols[3]
  bars <- symbols %in% c("B", "BB", "BBB")
  
  # get prize
  if (same) {
    payouts <- c("DD" = 100, "7" = 80, "BBB" = 40, "BB" = 25, 
      "B" = 10, "C" = 10, "0" = 0)
    prize <- unname(payouts[symbols[1]])
  } else if (all(bars)) {
    prize <- 5
  } else {
    cherries <- sum(symbols == "C")
    prize <- c(0, 2, 5)[cherries + 1]
  }
  

  # adjust for diamonds
  diamonds <- sum(symbols == "DD")
  prize * 2 ^ diamonds
}

score()

```



```{r}



df <- read_csv(here::here("gitlog"),col_names = FALSE)
  

df %>% 
  mutate(
    row_id=row_number()
    ,group=if_else(row_id %% 4==0,row_id/4,NA_integer_)
    ) %>% 
  fill(group,.direction = "up") %>% 
  mutate(commit_name=
           case_when(
             row_id %%4==0~ X1
            ,TRUE ~ NA_character_
           )
         )
  
```

### how to pass on functions within a function

call() and expr() are to simiiliar functions

call is limited to one function where expr can have multiple functions

```{r}
library(rlang)
library(tidyverse)


# how to create a call object
x <- call("mean",1:10)

y <- expr(mean(1:10))

z <- expr(mean(1:10))


## how to excute?
eval(x)

eval(y)

eval(!!!Z)


```


Turn an an expression back to symbol with rlang::as_string



```{r}
x <- expr(x)

rlang::as_string(x)
```

### how to use expr

- when you save something to expr it acts as a list
-   first subbetting returns the call
-   second returns the call arguments (use [[3]] or $name)


### visualizing expression
```{r}


lobstr::ast(
diamonds %>% 
  count(cut,sort=TRUE,wt=price)
  
  
)

```


```{r}
lobstr::ast(
f(
  g(
    h(1+2+3)
    ,(x+y)
    )
  ,z
  )
)
```


- the below there are related:
    
    -   enquo
    -   ensym -- this takes unquoted input and returns an expression. similiar to enexpr (which itself simliar to enquo) but thro errow when expressions are not symbols
    -   expr returns a express and works when arguments are directly inter to it
        -    however when used to caputre a formlaa arguments it wont work so you need enexpr to capture a user's arguments (vs. directly input in teh formual body)
    -   enexprs or exprs is used for multiple inputs seperate by ; returns list
    -ensym or ensym are like expr but checks is symbol or string
"" and !! are opposites

```{r}
cement <- function(...) {
  args <- ensyms(...)
  paste(purrr::map(args, as_string), collapse = " ")
}

time <- "test"

cement(!!time, hello)

outer <- 1:2


match.fun("*",outer)

match.fun()

match.fun("*")

outer <- 1:2
try(match.fun(outer, descend = FALSE)) #-> Error:  not a function
call("*",outer) %>% eval()



```
```{r}




```

