---
title: "interactions"
---

```{r}
library(tidyverse)


df <- fpaR::contoso_fact_sales %>% 
  janitor::clean_names() %>% 
  left_join(fpaR::contoso_dim_product %>% janitor::clean_names(),by = join_by(product_key))


df_summary <- df %>% 
  mutate(rev=unit_price.x*sales_quantity) %>% 
  summarise(
    rev=sum(rev)
    ,n=n()
    ,avg_vol=mean(sales_quantity)
    ,distinct_col=length(unique(color_name))
    ,.by = c(date_key,brand_name,color_name)
  )


df %>% count(manufacturer)

m1 <- df_summary %>% 
  lm(rev~brand_name+avg_vol
     ,data=.
  )

m2 <- df_summary %>% 
  lm(rev~brand_name*avg_vol
     ,data=.
  )

emmeans::emmeans(m2, pairwise ~brand_name|avg_vol,at=list(avg_vol=c(12,15,18)))
aov(m1,m2)

daily_summary <- df_summary %>% 
  summarise(daily_rev=sum(rev)
            ,mean_rev=mean(rev)
            ,prop_50=quantile(rev,.5)
            ,avg_vol=mean(avg_vol)
            ,n=n()
            ,.by=c(brand_name,color_name)
            )
```

```{r}

library(scales)
m1 <- df_summary %>% 
  lm(rev~0+brand_name*color_name,data=.)

lm_summary <- m1 %>% broom::tidy() %>%   mutate(term=str_remove(term,"brand_name|color_name"))
rq_summary <- df_summary %>% quantreg::rq(rev~0+brand_name+avg_vol,data=.) %>% broom::tidy(se="iid") %>%   mutate(term=str_remove(term,"brand_name")) %>% select(term,median=estimate)

m2 <- df_summary %>% 
  lm(rev~brand_name*avg_vol
     ,data=.
  )

daily_summary %>% 
  left_join(lm_summary,by = join_by(brand_name==term)) %>% 
  left_join(rq_summary,by = join_by(brand_name==term)) %>% 
  select(brand_name,daily_rev,estimate,mean_rev,prop_50,n,median) %>% 
  mutate(
    total_rev=n*estimate
    ,delta_mean=estimate-mean_rev
    ,delta_rev=daily_rev-total_rev
    )
lm_summary 

daily_summary <- daily_summary %>% 
  mutate(vol_effect=avg_vol*502.2374)


```


```{r}
lm_summary %>% 
  filter(str_detect(term,"Contoso:color_nameGrey|Contoso")) %>% pull(estimate) %>% sum(.,na.rm=TRUE)
```


### linear modeling for understanding business analysis

-   If want to to get the total and average revue for each brand, we can do summarize and use sum and mean

```{r}

daily_summary <- df_summary %>% 
  summarise(daily_rev=sum(rev)
            ,mean_rev=mean(rev)
            ,prop_50=quantile(rev,.5)
            ,avg_vol=mean(avg_vol)
            ,n=n()
            ,.by=brand_name)

daily_summary %>% 
  select(1:2,mean_rev)
```



-   We can replicate these results with a linear model by putting the thing we want to get the average of as y and then put the single dimension as x and forcing the intercept to be 0
```{r}
m1 <-   df_summary %>% lm(rev~0+brand_name,data=.)

lm_summary <- m1 %>% broom::tidy() %>%   mutate(term=str_remove(term,"brand_name"))

lm_summary
```


- the same can be replicated for median value and qr()


- if we add in two effects `lm(rev~0+brand_name+avg_vol,data=.)` then we get new estimates but no longer do we get the average by brand_name, instead we get the averge by brand_name after removing a fixed effect of avg_vol. In other words, if we take the average of that variable, multiply it by the esitmate, we will get the get revenue contribute due to avg_vol, the brand_name estimate average is being held contanst at the level
