---
title: "interactions"
---

```{r}

#load libraries and summarize data to be be used in later analysis
library(tidyverse)
devtools::install_github("alejandrohagan/fpaR")


df <- fpaR::contoso_fact_sales %>% 
  janitor::clean_names() %>% 
  left_join(fpaR::contoso_dim_product %>% janitor::clean_names(),by = join_by(product_key))


df_summary <- df %>% 
  mutate(rev=unit_price.x*sales_quantity) %>% 
  summarise(
    rev=sum(rev)
    ,n=n()
    ,vol=sum(sales_quantity)
    ,avg_vol=mean(sales_quantity)
    ,distinct_col=length(unique(color_name))
    ,.by = c(date_key,brand_name,color_name)
  ) %>% 
  mutate(vol_group=cut(vol,breaks = 5))

```



# First  comparison

-   When you do zero coefficent regression with a category variable it is the same as doing a group by and summarize
        
        -   This doesn't change if you emmeans or directly pull the case weights
        -
        

```{r}
#| echo: false
#| label: category-model
#| eval: TRUE

# zero intercept
category_lm_zero <- df_summary %>% 
  lm(rev~0+brand_name
     ,data=.
  )

# classic lm
category_lm<- df_summary %>% 
  lm(rev~brand_name
     ,data=.
  )


#extract out marginal means and weights

em_cat_lm_zero_tbl<- emmeans::emmeans(category_lm_zero, pairwise ~brand_name)
em_cat_lm_tbl<- emmeans::emmeans(category_lm, pairwise ~brand_name)

#do simple mean by group

cat_manual_tbl <- df_summary %>% 
  group_by(brand_name) %>% 
  summarize(avg_rev=mean(rev))

## join results
em_cat_lm_zero_tbl$emmeans %>% as_tibble() %>% 
  select(
    1
    ,emmean_cat_lm_zero=2) %>% 
  left_join(cat_manual_tbl,by=join_by(brand_name)) %>% 
  left_join(em_cat_lm_tbl$emmeans %>% as_tibble() %>% select(1,cat_lm=2)
            ,join_by(brand_name)) %>% 
  left_join(
    category_lm_zero %>% broom::tidy() %>% select(1,lm_weights=2,p.value) %>% mutate(term=str_remove(term,"brand_name"))
    ,by=join_by(brand_name==term)
  )

```

-emmeans and linear model and summarize are all equal when it is a single quantified model


```{r}

m2 <- df_summary %>% 
  lm(rev~brand_name*color_name
     ,na.action = NULL
     ,model = TRUE
     ,
     ,data=.
  )

em_tbl_2 <- emmeans::emmeans(m2, pairwise ~brand_name|color_name)

em_tbl <- em_tbl_2$emmeans %>% 
  as_tibble() %>% 
    mutate(term_name=paste0("brand_name",brand_name,":","color_name",color_name))
  

df_summary %>% 
  group_by(brand_name,color_name) %>% 
  summarize(avg_rev=mean(rev)
            ,n=n()) %>% 
  mutate(term_name=paste0("brand_name",brand_name,":","color_name",color_name)) %>% 
  left_join(em_tbl %>% select(term_name,emmean),
            by=join_by(term_name)) %>% 
  left_join(
    m2 %>% broom::tidy() %>% arrange(desc(p.value)) %>% select(term_name=term,estimate,p.value)
            ,by=join_by(term_name)
    )

 df_summary %>% 
  lm(rev~0+brand_name*color_name
     ,data=.
  ) %>% broom::tidy()

 
 df_summary$color_name %>% as_factor() %>% levels()
 
 37622.1988
 brand_nameContoso:color_nameWhite
 
 77251.5681
```

Interacting two variables is the same as grouping by these two variables and finding the average of that variable by that group. It's showing for that effect all else held equal (in the above case nothing is is needed) at least when passed thorugh emmeans -- when just passed as a linear model the estimates are different. Need to understand how emmeans vs. estimate are different

```{r}
emmeans::emmeans(m4, pairwise ~avg_vol,at=list(avg_vol=c(8,12,13.2,11,15,18,24)))
aov(m1,m2)

daily_summary <- df_summary %>% 
  summarise(daily_rev=sum(rev)
            ,mean_rev=mean(rev)
            ,prop_50=quantile(rev,.5)
            ,avg_vol=mean(avg_vol)
            ,n=n()
            ,.by=c(brand_name,color_name)
            )
```

- do  a simple model just vol

- now do a vol by interaction with a brand

- compare against summary and see what comes up 

- compare with zero constraint and without zero constraint




```{r}

library(scales)
m1 <- df_summary %>% 
  lm(rev~0+brand_name*color_name,data=.)

lm_summary <- m1 %>% broom::tidy() %>%   mutate(term=str_remove(term,"brand_name|color_name"))
rq_summary <- df_summary %>% quantreg::rq(rev~0+brand_name+avg_vol,data=.) %>% broom::tidy(se="iid") %>%   mutate(term=str_remove(term,"brand_name")) %>% select(term,median=estimate)

m2 <- df_summary %>% 
  lm(rev~brand_name*avg_vol
     ,data=.
  )

daily_summary %>% 
  left_join(lm_summary,by = join_by(brand_name==term)) %>% 
  left_join(rq_summary,by = join_by(brand_name==term)) %>% 
  select(brand_name,daily_rev,estimate,mean_rev,prop_50,n,median) %>% 
  mutate(
    total_rev=n*estimate
    ,delta_mean=estimate-mean_rev
    ,delta_rev=daily_rev-total_rev
    )
lm_summary 

daily_summary <- daily_summary %>% 
  mutate(vol_effect=avg_vol*502.2374)


```


```{r}
lm_summary %>% 
  filter(str_detect(term,"Contoso:color_nameGrey|Contoso")) %>% pull(estimate) %>% sum(.,na.rm=TRUE)
```


### linear modeling for understanding business analysis

-   If want to to get the total and average revue for each brand, we can do summarize and use sum and mean

```{r}

daily_summary <- df_summary %>% 
  summarise(daily_rev=sum(rev)
            ,mean_rev=mean(rev)
            ,prop_50=quantile(rev,.5)
            ,avg_vol=mean(avg_vol)
            ,n=n()
            ,.by=brand_name)

daily_summary %>% 
  select(1:2,mean_rev)
```



-   We can replicate these results with a linear model by putting the thing we want to get the average of as y and then put the single dimension as x and forcing the intercept to be 0
```{r}
m1 <-   df_summary %>% lm(rev~0+brand_name,data=.)

lm_summary <- m1 %>% broom::tidy() %>%   mutate(term=str_remove(term,"brand_name"))

lm_summary
```


- the same can be replicated for median value and qr()


- if we add in two effects `lm(rev~0+brand_name+avg_vol,data=.)` then we get new estimates but no longer do we get the average by brand_name, instead we get the averge by brand_name after removing a fixed effect of avg_vol. In other words, if we take the average of that variable, multiply it by the esitmate, we will get the get revenue contribute due to avg_vol, the brand_name estimate average is being held contanst at the level
