---
title: "interactions"
---

```{r}
#load libraries and summarize data to be be used in later analysis
library(tidyverse)
devtools::install_github("alejandrohagan/fpaR")


df <- fpaR::contoso_fact_sales %>% 
  janitor::clean_names() %>% 
  left_join(fpaR::contoso_dim_product %>% janitor::clean_names(),by = join_by(product_key))


df_summary <- df %>% 
  mutate(rev=unit_price.x*sales_quantity) %>% 
  summarise(
    rev=sum(rev)
    ,n=n()
    ,vol=sum(sales_quantity)
    ,avg_vol=mean(sales_quantity)
    ,distinct_col=length(unique(color_name))
    ,.by = c(date_key,brand_name,color_name)
  ) %>% 
  mutate(vol_group=cut(vol,breaks = 5))

```



```{r}



compare_simple_categorical_only <- function(.data,dependent,independent) {

dependent=dependent
independent=str_flatten(string = independent,collapse = " * ")

remove_cols <- str_flatten(string=independent,collapse="|")


zero_formula <- glue::glue(dependent," ~ 0 + ",independent)
  

reg_formula <- glue::glue(dependent," ~ ",independent)

reg_expression <- expression(reg_formula)
  
  category_lm <-lm(eval(zero_formula),data=.data) %>% broom::tidy() %>% mutate(term_clean=str_remove(term,remove_cols))
  
  category_lm_zero <- lm(eval(reg_formula),data=.data) %>% broom::tidy() %>% mutate(term_clean=str_remove(term,remove_cols))

# em_cat_lm_zero_tbl<- emmeans::emmeans(category_lm_zero, pairwise ~brand_name)
# em_cat_lm_tbl<- emmeans::emmeans(category_lm, pairwise ~brand_name)

  cat_manual_tbl <- .data %>% 
  group_by(any_of(independent)) %>% 
  summarize(avg_rev=mean(rev,na.rm=TRUE))
  
  out <- left_join(cat_manual_tbl)

  
  return(out)
  
}

remove_me <- "brand_name|color_name"
remove_me2 <- "color_name"
lm(rev~brand_name*color_name,df_summary) %>% 
  broom::tidy() %>% 
  mutate(term_clean=str_remove(term,pattern=remove_me))

df_summary %>% 
  group_by(brand_name,color_name) %>% 
  summarize(avg_rev=mean(rev,na.rm=TRUE)) 
  unite(col=term_clean,"brand_name","color_name")

```



# one categorical variable

-   When you do zero coefficent regression with a category variable it is the same as doing a group by and summarize
        
        -   This doesn't change if you emmeans or directly pull the case weights
        

```{r}
#| echo: false
#| label: category-model
#| eval: TRUE

# zero intercept
category_lm_zero <- df_summary %>% 
  lm(rev ~ 0+brand_name,data=.) %>% broom::tidy()
# classic lm
category_lm<- df_summary %>% 
    lm(rev ~ brand_name,data=.) %>% broom::tidy()


#extract out marginal means and weights

em_cat_lm_zero_tbl<- emmeans::emmeans(category_lm_zero, pairwise ~brand_name)
em_cat_lm_tbl<- emmeans::emmeans(category_lm, pairwise ~brand_name)

#do simple mean by group

cat_manual_tbl <- df_summary %>% 
  group_by(brand_name) %>% 
  summarize(avg_rev=mean(rev))

## join results

category_lm_tbl <- category_lm %>% mutate(term_clean=str_remove(term,"brand_name")) %>% select(term_clean,lm_intercept=estimate)
category_lm_zero_tbl <- category_lm_zero %>%  mutate(term_clean=str_remove(term,"brand_name")) %>% select(term_clean,lm_zero=estimate)


summary_results <- cat_manual_tbl %>% 
  left_join(
    category_lm_tbl
    ,by=join_by(brand_name==term_clean)
  ) %>% 
  left_join(
    category_lm_zero_tbl
    ,by=join_by(brand_name==term_clean)
  )

```

-emmeans and linear model and summarize are all equal when it is a single quantified model

## Two categorical variables

```{r}
#| echo: false
#| label: category-model
#| eval: TRUE

# zero intercept
category_lm_zero <- df_summary %>% 
  lm(rev ~ 0+brand_name*color_name,data=.) %>% broom::tidy()
# classic lm
category_lm<- df_summary %>% 
    lm(rev ~ brand_name*color_name,data=.) %>% broom::tidy()


#extract out marginal means and weights

em_cat_lm_zero_tbl<- emmeans::emmeans(category_lm_zero, pairwise ~brand_name)
em_cat_lm_tbl<- emmeans::emmeans(category_lm, pairwise ~brand_name)

#do simple mean by group

cat_manual_tbl <- df_summary %>% 
  group_by(brand_name,color_name) %>% 
  summarize(avg_rev=mean(rev)) %>% 
  ungroup() %>% 
  mutate(term_clean=paste0(brand_name,":","color_name",color_name)) %>% 
  select(term_clean,avg_rev)

## join results

category_lm_tbl <- category_lm %>% mutate(term_clean=str_remove(term,"brand_name|color_name")) %>% select(term_clean,lm_intercept=estimate)
category_lm_zero_tbl <- category_lm_zero %>%  mutate(term_clean=str_remove(term,"brand_name|color_name")) %>% select(term_clean,lm_zero=estimate)


summary_results <- cat_manual_tbl %>% 
  left_join(
    category_lm_tbl
    ,by=join_by(term_clean==term_clean)
  ) %>% 
  left_join(
    category_lm_zero_tbl
    ,by=join_by(term_clean==term_clean)
  )

```


```{r}
lm_summary %>% 
  filter(str_detect(term,"Contoso:color_nameGrey|Contoso")) %>% pull(estimate) %>% sum(.,na.rm=TRUE)
```


### linear modeling for understanding business analysis

-   If want to to get the total and average revue for each brand, we can do summarize and use sum and mean

```{r}

daily_summary <- df_summary %>% 
  summarise(daily_rev=sum(rev)
            ,mean_rev=mean(rev)
            ,prop_50=quantile(rev,.5)
            ,avg_vol=mean(avg_vol)
            ,n=n()
            ,.by=brand_name)

daily_summary %>% 
  select(1:2,mean_rev)
```



-   We can replicate these results with a linear model by putting the thing we want to get the average of as y and then put the single dimension as x and forcing the intercept to be 0
```{r}
m1 <-   df_summary %>% lm(rev~0+brand_name,data=.)

lm_summary <- m1 %>% broom::tidy() %>%   mutate(term=str_remove(term,"brand_name"))

lm_summary
```


- the same can be replicated for median value and qr()


- if we add in two effects `lm(rev~0+brand_name+avg_vol,data=.)` then we get new estimates but no longer do we get the average by brand_name, instead we get the averge by brand_name after removing a fixed effect of avg_vol. In other words, if we take the average of that variable, multiply it by the esitmate, we will get the get revenue contribute due to avg_vol, the brand_name estimate average is being held contanst at the level
