---
title: "tidymodel_practice"
format: html
editor: visual
---

## Required packages

-   Tidymodels
    -   `rsample` used to split your sample
    -   `recipes` used to create recipe and transform data (normlize)
    -   `parsnip` used to create the model, engine and mode
    -   `workflow` puts together model (parsnip) and formula (recipe)
    -   `parsnip` also fits the everything together
-   usemodels can help set up your workflow including tunning and grid
-   There is no "set" pattern and these steps can be combined through other variables however below is a typical workflow

### Load libraries

```{r}
#| echo: true
pacman::p_load(tidyverse,scales,tidymodels,janitor,lubridate,vip)
```

### rsample

-   first rsample::initial_split(df,strata=x) to split your data set into training and testing
    -   `strata` ensures the predicated variable is kept constant in both samples
-   from here, you need use that split object with `training_split()` and `testing_split()`
-   this will create a training data set and testing data set (respectively)
    -   its possible to further break up your training data set into a validation set
-   you can further "split" your training data into either folds or bootstrap
-   `rsample::vfold_cv()` is fold your data into sub samples
    -   `strata` arguments makes sure the outcome variables is equally represented in each sample
    -   `v` is the number of times "fold" the data, default is 10
    -   `repeats` will repeat the folding patterns x times, default is 1

```{r split-data}
#| echo: true
#| eval: true


diamonds <- diamonds %>% 
  mutate(high_price_indicator=if_else(price>10000,"high","low"),
         high_price_indicator=factor(high_price_indicator),
         across(c("cut","color","clarity"),~as.factor(.)))


data_split <- rsample::initial_split(diamonds,strata = high_price_indicator)

training_split <- rsample::training(data_split)
testing_split <- rsample::testing(data_split)


diamonds_fold <- rsample::vfold_cv(training_split,strata=high_price_indicator)

```

## Parsnip

-This is package where you choose the model and configure it (set its engine and mode) -most models start with their modeling name (eg.`logistic_reg()`, `rand_forest()`, `linear_reg()`) - following selection the model you need select their engine (typical - name of the underlying package) and their mode (not always applicable; eg classification or regression) - most models have default modes and engine

```{r choose-model}
#| echo: true
#| eval: true
#| warning: false

log_spec <- parsnip::logistic_reg() %>% 
  parsnip::set_mode("classification") %>%
  parsnip::set_engine("glm")



rf_spec <- parsnip::rand_forest(trees = 1000) %>% 
  set_mode("classification") %>% 
  set_engine("ranger")

```

### Recipes

-   this is where you put the formula and transformation steps that you want your model
-   typical to start with recipes::recipe(outcome \~ predictor1+ predictor2, data=df_training)
-   you can then transform data with `step_?` using tidy select principles
    -   for the dataset you should put your training set (not your folded data set or testing set)
    -   `step_other()` takes high variable data and categories minor categories into other
    -   `step_normalize()` takes data and scales and center its - either directly reference or use `all_numeric()` as a shortcut
    -   `step_dummy()` which transform categorical variable into columns, use `one_hot=TRUE` ; variables should all be factors
    -   `step_zv()` to remove variables with zero variation (can use with `all_predictors()`)
-   while you may have many models and parameters, you would typically have one receipe that can be leverage many times


- the dataset only gets used to help recipes determine the column types
- [recipes links](https://www.tmwr.org/recipes.html)   


### prep & bake

-applies recipes steps to a dataset
- prep(rec,new_data=null)

[prep_and_bake](https://www.tmwr.org/dimensionality.html)


```{r}
rec <- recipes::recipe(high_price_indicator ~., data=training_split) %>%
  step_string2factor(c("cut","color","clarity")) %>% 
  recipes::step_normalize(all_numeric_predictors()) %>% 
  step_zv(all_predictors()) %>% 
  step_dummy(c("cut","color","clarity"),one_hot = TRUE) %>% 
  step_rm(price)


prep(rec,training_split) %>% juice()

```

### Workflow

-   this stitches together your various elements (model, recipe)
-   I need to learn more about this but appears it makes it eaier to do thigns later one -add_model(model_spec) -add_recipe(recipe_spec)
-   You can pass multiple reciepes or models via a list to a workflow to `workflow_set(preproc=recipe_list,models=model_list)`
-   extract specific models `extact_workflow(workflow_set,id=workflow_id)`
-   Mainly designed to be used with resamped data

common formulas - `workflow()` - `workflow_set()` if passing multiple formlas or models

::: callout-note
from the learning site, workflows section of [tidy models book](https://www.tmwr.org/workflows.html), workflows are : if you are using multiple model types, or constantly updating various elements of the model then workflow allows to add / substrat specific elements (without rebuilding the entire model from scratch)

This "full model" is compared to a sequence of the same model that removes each predictor in turn. Using basic hypothesis testing methods or empirical validation, the effect of each predictor can be isolated and assessed.
:::

link:

```{r}
workflow <- workflow() %>% 
  add_model(rf_spec) %>% 
  add_recipe(rec)

```

### Fit Model

-   This is when you actually apply fit the model to your data and do the "machine learning"

-   if you did not do the `rsample::vfold_cv()` step then you use the `parsnip::fit()`

    -   parsnip::fit(df, workflow)

-   if you did use `rsample::vfold_cv()` then use `tune::fit_resamples`

-   `workflow()` - insert your workflow here

-   `resamples= folds` - this is where you set the folds (data)

-   `control = control_resamples(save_pred = TRUE)` -need to manually save predictions

-   this can take alot of time so turn on parellel processing `doParallel::registerDoParallel()`

```{r fit}


doParallel::registerDoParallel()

### if you did not vfold_cv()

#fit_rf <- workflow %>%
 # fit(data=training_split)


### if you did vfold
fit_folds <- tune::fit_resamples(workflow,
    resamples = diamonds_fold,
    control = control_resamples(save_pred = TRUE)
  )


```

Validate model - if you want to see your samples you can do the `collect_metrics(rs_data)`

```{r}
collect_metrics(fit_folds)
```

-   use the roc_curve to plot your results - if resampled group by the resampled id
-   the variable predicted probabilities will use the `pred_var1` and `pred_var1` name

```{r}
#| eval: false
collect_predictions(fit_folds) %>%
  group_by(id) %>%
  roc_curve(high_price_indicator,.pred_high) %>%
  autoplot()
```

### confusion matrix on training data

-   can also do a confusion matrix - no need to unnest data

```{r}
conf_mat_resampled(fit_folds, tidy = FALSE) %>%
  autoplot()


```

-   you can also apply the confusion matrix on the entire set (including testing set)
-   must use `last_fit(workfow, original_split)` -- I don't super undersand why
-   form there follow same path

```{r}
final_fitted <- last_fit(workflow, data_split)

collect_metrics(final_fitted)

collect_predictions(final_fitted) %>%
  conf_mat(high_price_indicator, .pred_class) %>%
  autoplot()

collect_predictions(final_fitted) %>%
  roc_curve(high_price_indicator,.pred_high) %>%
  autoplot()

```

-   Can also rank the feature importance with VIP packge
-   take recipe, prep it and apply it blank data -- still dont under prep() vs. juice() vs. bake()
-   

```{r}
library(vip)

imp_data <- rec %>%
  prep() %>%
  bake(new_data = NULL)

rf_spec %>%
  set_engine("ranger", importance = "permutation") %>%
  fit(high_price_indicator ~ ., data = imp_data) %>%
  vip(geom = "point")
```

### Random Forest

-   okay to use between 500 - 1000 trees
-   can set your cores directly in rand_forest() argument (cores \<- `parallel::detectCores()`)
-   turn predicted varible into factor

#### Shortcuts to remember things

Great shortcut to set up your model so that you don't have to remember everything - you need to know your receipe and training split ahead of time

##### Usemodels





```{r}


df <- tibble(x=1:4,
       y=letters[1:4],
       z=c("alpha","omega","charlie","echo"))

rec_2 <- recipes::recipe(x~.,data=df) %>% 
  step_dummy(z,one_hot = TRUE)

rec_2 %>% prep() %>% juice()
```

