---
title: "tidymodel_practice"
format: html
--- 



## Required packages

-   Tidymodels
    -   `rsample` used to split your sample
    -   `recipes` used to create recipe and transform data (normlize)
    -   `parsnip` used to create the model, engine and mode
    -   `workflow` puts together model (parsnip) and formula (recipe)
    -   `parsnip` also fits the everything together
-   usemodels can help set up your workflow including tunning and grid
-   There is no "set" pattern and these steps can be combined through other variables however below is a typical workflow

### Load libraries

```{r}
#| echo: true
pacman::p_load(tidyverse,scales,tidymodels,janitor,lubridate,vip)
```

### rsample

-   First rsample::initial_split(df,strata=x,prop=y) to split your data set into training and testing and assign it to an object
    
    -   `strata` ensures the predicated variable is kept constant in both samples
    -   'prop' determins the proportion of of initial vs. split   
-   From here, you need use that split object with `training_split()` and `testing_split()` and also assign to new objects
-   This will create a training data set and testing data set (respectively)

    -   its possible to further break up your training data set into a validation set

-   you can further "split" your training data into either folds or bootstrap
-   `rsample::vfold_cv()` is fold your data into sub samples

    -   `strata` arguments makes sure the outcome variables is equally represented in each sample
    -   `v` is the number of times "fold" the data, default is 10
    -   `repeats` will repeat the folding patterns x times, default is 1

```{r split-data}
#| echo: true
#| eval: true


diamonds <- diamonds %>% 
  mutate(high_price_indicator=if_else(price>10000,"high","low"),
         high_price_indicator=factor(high_price_indicator),
         across(c("cut","color","clarity"),~as.factor(.)))


data_split <- rsample::initial_split(diamonds,strata = high_price_indicator)

training_split <- rsample::training(data_split)
testing_split <- rsample::testing(data_split)


diamonds_fold <- rsample::vfold_cv(training_split,strata=high_price_indicator)

```

## Parsnip

- This is package where you choose the model, engeine and mode and configure it (set its engine and mode)
-   most models start with their modeling name (eg.`logistic_reg()`, `rand_forest()`, `linear_reg()`) 
- Following selection the model you need select their engine (typical - name of the underlying package) and their mode (not always applicable; eg classification or regression) - most models have default modes and engine
-   First start with the model eg. linear_reg

    -   Then choose the mode arugment in the model(mode="regression"), for some there is an option (regression or classification), for others it can only be one so you can't change it
-   Then set the engine (set_egine()) which determines which underlying package you want to use (eg. stats::lm, spark, etc)
-   optional: At this point, you have the option of just directly appliying parsnip::fit() into the model by directly entering recipe and then data
    
    -   fit(model,price ~ category, data=train_tbl) and this will directly  output results
    -   if you do this you cna directly put in `predict(model_fit, new_data=test_data)` 


```{r choose-model}
#| echo: true
#| eval: true
#| warning: false

log_spec <- parsnip::logistic_reg() %>% 
  parsnip::set_mode("classification") %>%
  parsnip::set_engine("glm")



rf_spec <- parsnip::rand_forest(trees = 1000) %>% 
  set_mode("classification") %>% 
  set_engine("ranger")


rf_glmnet <- parsnip::linear_reg(penalty = 10,mixture = 0.5) %>% 
  parnship::set_mod("regression") %>% 
  parsnip::set_engine("glmnet")


rf_decision <- decision_tree(cost_complexity = 0.01,min_n=10) %>% 
  set_mode("regression") %>% 
  set_engine("rpart")


```

### Recipes

-   this is where you put the formula and transformation steps that you want your model

-   typical to start with recipes::recipe(outcome \~ predictor1+ predictor2, data=df_training)

-   you can then transform data with `step_?` using tidy select principles

    -   for the dataset you should put your training set (not your folded data set or testing set)
    -   `step_other()` takes high variable data and categories minor categories into other
    -   `step_normalize()` takes data and scales and center its - either directly reference or use `all_numeric()` as a shortcut
    -   `step_dummy()` which transform categorical variable into columns, use `one_hot=TRUE` ; variables should all be factors
    -   `step_zv()` to remove variables with zero variation (can use with `all_predictors()`)

-   while you may have many models and parameters, you would typically have one receipe that can be leverage many times

-   the dataset only gets used to help recipes determine the column types

-   [recipes links](https://www.tmwr.org/recipes.html)

### prep & bake

-applies recipes steps to a dataset - prep(rec,new_data=null)

[prep_and_bake](https://www.tmwr.org/dimensionality.html)

```{r}
rec <- recipes::recipe(high_price_indicator ~., data=training_split) %>%
  step_string2factor(c("cut","color","clarity")) %>% 
  recipes::step_normalize(all_numeric_predictors()) %>% 
  step_zv(all_predictors()) %>% 
  step_dummy(c("cut","color","clarity"),one_hot = TRUE) %>% 
  step_rm(price)


prep(rec,training_split) %>% juice()

```

### Workflow

-   this stitches together your various elements (model, recipe)
-   I need to learn more about this but appears it makes it eaier to do thigns later one -add_model(model_spec) -add_recipe(recipe_spec)
-   You can pass multiple reciepes or models via a list to a workflow to `workflow_set(preproc=recipe_list,models=model_list)`
-   extract specific models `extact_workflow(workflow_set,id=workflow_id)`
-   Mainly designed to be used with resamped data

common formulas - `workflow()` - `workflow_set()` if passing multiple formlas or models

::: callout-note
from the learning site, workflows section of [tidy models book](https://www.tmwr.org/workflows.html), workflows are : if you are using multiple model types, or constantly updating various elements of the model then workflow allows to add / substrat specific elements (without rebuilding the entire model from scratch)

This "full model" is compared to a sequence of the same model that removes each predictor in turn. Using basic hypothesis testing methods or empirical validation, the effect of each predictor can be isolated and assessed.
:::

link:

```{r}
workflow <- workflow() %>% 
  add_model(rf_spec) %>% 
  add_recipe(rec)

```

### Fit Model

-   This is when you actually apply fit the model to your data and do the "machine learning"

-   if you did not do the `rsample::vfold_cv()` step then you use the `parsnip::fit()`

    -   parsnip::fit(df, workflow)

-   if you did use `rsample::vfold_cv()` then use `tune::fit_resamples`

-   `workflow()` - insert your workflow here

-   `resamples= folds` - this is where you set the folds (data)

-   `control = control_resamples(save_pred = TRUE)` -need to manually save predictions

-   this can take alot of time so turn on parellel processing `doParallel::registerDoParallel()`

```{r fit}


doParallel::registerDoParallel()

### if you did not vfold_cv()

#fit_rf <- workflow %>%
 # fit(data=training_split)


### if you did vfold
fit_folds <- tune::fit_resamples(workflow,
    resamples = diamonds_fold,
    control = control_resamples(save_pred = TRUE)
  )


```


## Yardsticks

- once you ahve a tbl with test results and prediction columns
- easiest way is to use `yardstricks::metrics()` or `yardstick::calc_metrics()`
  -   truth= what you are testing for
  -   estimate= what you predicted
- this will product standard metrics like mae,rmse and rsq


Validate model - if you want to see your samples you can do the `collect_metrics(rs_data)`

```{r}
collect_metrics(fit_folds)
```

-   use the roc_curve to plot your results - if resampled group by the resampled id
-   the variable predicted probabilities will use the `pred_var1` and `pred_var1` name

```{r}
#| eval: false
collect_predictions(fit_folds) %>%
  group_by(id) %>%
  roc_curve(high_price_indicator,.pred_high) %>%
  autoplot()
```

### confusion matrix on training data

-   can also do a confusion matrix - no need to unnest data

```{r}
conf_mat_resampled(fit_folds, tidy = FALSE) %>%
  autoplot()


```

-   you can also apply the confusion matrix on the entire set (including testing set)
-   must use `last_fit(workfow, original_split)` -- I don't super undersand why
-   form there follow same path

```{r}
final_fitted <- last_fit(workflow, data_split)

collect_metrics(final_fitted)

collect_predictions(final_fitted) %>%
  conf_mat(high_price_indicator, .pred_class) %>%
  autoplot()

collect_predictions(final_fitted) %>%
  roc_curve(high_price_indicator,.pred_high) %>%
  autoplot()

```

-   Can also rank the feature importance with VIP packge
-   take recipe, prep it and apply it blank data -- still dont under prep() vs. juice() vs. bake()
-   

```{r}
library(vip)

imp_data <- rec %>%
  prep() %>%
  bake(new_data = NULL)

rf_spec %>%
  set_engine("ranger", importance = "permutation") %>%
  fit(high_price_indicator ~ ., data = imp_data) %>%
  vip(geom = "point")
```

### Random Forest

-   okay to use between 500 - 1000 trees
-   can set your cores directly in rand_forest() argument (cores \<- `parallel::detectCores()`)
-   turn predicted varible into factor

#### Shortcuts to remember things

Great shortcut to set up your model so that you don't have to remember everything - you need to know your receipe and training split ahead of time

##### Usemodels

```{r}


df <- tibble(x=1:4,
       y=letters[1:4],
       z=c("alpha","omega","charlie","echo"))

rec_2 <- recipes::recipe(x~.,data=df) %>% 
  step_dummy(z,one_hot = TRUE)

rec_2 %>% prep() %>% juice()
```
<<<<<<< HEAD
=======


## ploting reuslts

-    take the fit object and broom::tidy()
-   


## To be classified lateR:


#tidymodels

Split data & sample folds

1.  rsample::initial_sample(dataframe,strata=variable)

2.  rsample::training(split_data)

3.  rsample::testing(split_data)

4.  vfold_cv(df_testing, strata = type)

#### recipe to create formula and transform daata

1.  recipes::recpe(outcome \~ predictor,data=train_df)

#### select the ml model and type

parsnip::logistic_reg() %\>% set_mode() %\>% set_engine()

###workflow

workflow() %\>% add_recipe(rec) %\>% add_model(mod)





## kmeans



- step 1: 
organize  data into normalied format

whatever we tryign to study (eg. product purchases) then you fill in the percent of that produc tby the dimenion you want (customer), eg customer by product

- step 2:
use kmeans function with table and initial group selection

-only can have numeric input (so create with explantory varibale then deselect variable (eg. customers))

kmeans_obj = kmeans(nomarlized_tbl,
centers=num__of_groups,
iter.max=something,
nstarts= higher is better (say 100))



kmeans_obj$center = all various centers

kmeans_obj$cluster=classify each obs to a clsuter

step 3: apply kmeans_obj to data

broom::augment(kmeans_obj,normalized_tbl)

step4: how to intrpret
tot.withinss is metric of total squared varition (lower is better), 

broom::glance(kmeans_obj)]
g

-   step5 optimze by setting up a grid

create function that calcules kmeans with cneters (and nstarts if you want) to be iterated with an input table and pipe this into glance

- plot tot.withinss and cneters to see which centers is optimized

# kmeans example

```{r}
devtools::load_all("/home/hagan/R/fpaR")
sales_by_store_product <- fpaR::contoso_fact_sales %>% 
  janitor::clean_names() %>% 
  mutate(sales_amount=unit_price*sales_quantity) %>% 
  group_by(store_key,product_key) %>% 
  summarise(sum_usd=sum(sales_amount),.groups = "drop") %>% 
  arrange(product_key) %>% 
  group_by(store_key) %>% 
  mutate(prop_usd=sum_usd/sum(sum_usd)) %>% 
  select(store_key,product_key,prop_usd) %>% 
  pivot_wider(names_from=product_key,values_from=prop_usd,values_fill = 0,names_repair = janitor::make_clean_names)


kmeans_obj <- sales_by_store_product %>% 
  select(-store_key) %>% 
  kmeans(centers=15,nstart=100,iter.max = 100)

stores_with_kmeans <- kmeans_obj %>% broom::augment(data=sales_by_store_product) 
kmeans_obj %>% broom::glance()


kmeans_mapper <- function(centers=3) {
  sales_by_store_product %>% kmeans(x=.,centers=centers,nstart = 100,iter.max=100)
  
}

grid_kmeans <- tibble(centers=1:15) %>% 
  mutate(kmeans_obj=map(.x=centers,~kmeans_mapper(.x)),
         kmeans_output=map_df(kmeans_obj,broom::glance)
  )

grid_kmeans %>%
  unnest(kmeans_output) %>% 
  select(centers,tot.withinss) %>% 
  ggplot(aes(x=centers,y=tot.withinss))+
  geom_point()+
  geom_line()

```

## umap

very simliar to kmeans just without addition arguments
stregth is merge the kmeans output with the kmeans output so that you can mpa them together
then manually inspect original data frame by the new clusters to see what patterns or attributes (eg. can sort by some cumsum and attributes to look for patterns)


step1 
- use same normalized model
- only select numeric data
- pass through to umeans

step2
- take hte umeans$layout covert to tibble, rename columns and bind with framing column




```{r}

library(umap)
umap_obj <- sales_by_store_product %>% 
  select(-store_key) %>% 
  umap()

umap_tbl <- umap_obj$layout %>%
  as_tibble() %>% 
  set_names(c("x","y")) %>% 
  bind_cols(sales_by_store_product) %>% 
  select(store_key,x,y)

umap_tbl %>% 
  ggplot(aes(x,y))+
  geom_point()
  
kmeans_result_tbl <-   grid_kmeans %>%
    filter(centers==8) %>% 
    pull(kmeans_obj) %>% 
    pluck(1) %>% broom::augment(sales_by_store_product) %>% 
    relocate(last_col()) %>% 
    select(.cluster,store_key)


kmeans_result_tbl %>% 
  left_join(umap_tbl,by="store_key") %>% 
  ggplot(aes(x=x,y=y,col=.cluster))+
  geom_point()

```


## combine umeans and kmeans

>>>>>>> a548534aa2ba9c3ce865bab0e837a2867fc5c783
