# Chapter 1: Let's get started

# Named Objects

::: callout-note
In this package, we will use the following libraries

library(tidyverse)
:::

```{r}
#|echo: false

pacman::p_load(tidyverse)

```


### how to create and reference objects and generate series

Although not the most glamorous place to start, learning how to create named objects, reference and manipulate them as well generate data becomes incredibly useful down the road when combined with other techniques and workflows.

In R this is main way you save, reference, and manipulate your data

You use the `<-` (ctrl + - ) or the `=` symbol to assign a result to an object

```{r}


#| echo: true
object <- 1+1
```

Whatever name type will become the object's name, if type in the object, it will return whatever you assigned to the name

Although this is fairly intuitive to understand, if you are coming from an Excel background, it will take time to develop this habit of assigning names to results because in Excel, you can refer to a result by its cell location by physically pointing a formula to a input.

For quick and simple calculations, this will be quicker but if you try to do something even beyond a simple calculation, you will see it becomes very difficult to manage, easy to make mistakes and reduces your producivity as you have to constantly point to those input cells or have to remember non-intuitive names `eg C5` .

In excel, you can also name a cell by putting your cursor over the cell and click on the named box in the top left.

To keep track of all the names, you can press F5 and you will see a dialog box that will help you manage all of your names.

When you want to refer to that cell, you can simply type in the cell name.

Although useful for intermediate analysis, I would say from a workflow perspective, it isn't "easy" as you have to constantly break your workflow but it will make your work more readable.

### how to create and named Tables

##### In R

Regardless if the data object is a cell or a table (or actually any type of object) the naming methodology is the same.

However creating a table isn't as straight forward because you don't have the same GUI.

Data entry is not R (or any scripting languages) strong suit - which doesn't mean you can't do it but if you are strictly entering in data (of great length), use excel.

However, below shows you how to create tables

#### tibble / data frame

```{r}
library(tidyverse)
df_example <- tibble(x=1:10,
                     y=2:11,
                     z=3:12
                     )

```

#### Tribble

When trying to do data entry that you need to reference columns wise & rowwise, this can challenging, so thats why there is a `tribble`, a handy way to still see your columnwise and rowwise data together

::: callout-note
There is a concept of vectorization, which for now, we will define as recycling. Which is to say if I put one column with 10 elements, and another column with only 1 element, tibble will recycle the "1" until it length of the tibble.
:::

```{r}
df_tribble <- tribble(
  ~col1,~col2,~col3,
  1,2,3,
  5,6,7,
  8,9,10
  
  
)
```

Another methodology which can be useful are the `data.edit` and

```{r}

```

::: callout-tip
What naming convention should we use in R? This is where you will start to see namng conventions that you are not familiar with such as snake_case, kebab-case, or Upper_camel.

For this book, we will use snake_case (but you use whatever is easist for you).

Coming up with names for variables will be challenging as you will either type too little or too much, over time you will get your own convention. It is not important what is optimal, more important to focus on convention (means consistency) so that when you read something you immediately know what it is.
:::

### Excel Skills to be covered

Named cells 

Named Tables 

Adding columns to Tables 

Adding rows to Tables 

Referencing named objects 

F5 

Index() 

Tab references 

How to edit named objects 

Data Entry Form add in 

Creating vector 

Fill Series Tips 

Existing data 

New series of Numbers 

New series of Dates 

Flash Fill tips 

Adding/subtracting/dividing/multiplying columns 

One formula works against all rows in a column 

Rowwise equvalant 

Colwise equvailant 

Mean 

Median 

Sum 

Count 

Aggregate 

Error handling 

Iferror() 

NA vs. null vs. ref 

Array \[what did I mean by this\]

## R skills to be covered

Tibbles 

Lists 

Vectors 

dataeditR 

Arrange() 

Generating data 

-numbers 

-letters 

-dates 

-rep() 

-seq() 

-mutate

::: {#special .sidebar}
::: warning
Here is a warning.
:::

More content.
:::



`tibble::enframe()` takes a named list and turns into a tibble


I(rgb) will take the literal names f the columns and apply the colors https://www.youtube.com/watch?v=CTWkJrvfRBc minute 37:42

#### Metadata for columns

resource:

\[link to original source\](https://www.pipinghotdata.com/posts/2022-09-13-the-case-for-variable-labels-in-r/)

    penguins_metadata <- tribble(
        ~variable,           ~variable_label, 
        "species",           "Penguin species",
        "island",            "Island in Palmer Archipelago, Antarctica",
        "bill_length_mm",    "Bill length (mm)",
        "bill_depth_mm",     "Bill depth (mm)",
        "flipper_length_mm", "Flipper length (mm)",
        "body_mass_g",       "Body mass (g)",
        "sex",               "Penguin sex",
        "year",              "Study year"
    )

`labelled::generate_dictionary()`

generates a data dictionary

subsittute labels for points in graph

To substitute variable labels for variable names on your input variables, add [`ggeasy::easy_labs()`](https://jonocarroll.github.io/ggeasy/reference/easy_labs.html) to your ggplot.

    penguins_labelled <- penguins |> 
      set_variable_labels(
        species           = "Penguin species",
        island            = "Island in Palmer Archipelago, Antarctica",
        bill_length_mm    = "Bill length (mm)",
        bill_depth_mm     = "Bill depth (mm)",
        flipper_length_mm = "Flipper length (mm)",
        body_mass_g       = "Body mass (g)",
        sex               = "Penguin sex",
        year              = "Study year"
      )

    penguins_labelled <- penguins |> 
      set_variable_labels(!!!penguins_labels)



#### Quarto style guides

As best I can tell there are 3 ways to format presentations


`:::{.args input1="..." input2="...}`

impact text
`:::`
- this appears to reference built in quarto arguments
- they start with .args followed by inputs into that args

{style="args:..."}`
- this appears to reference CSS style arguments

##### Change changing style within string
If you want to edit a single word or string of  words **within a string** 


[existing text]{style="args:..."}


or if you want to edit a whole string (not just parts of it)

`:::{style="args:..."}` 
insert text
`:::` 

*popular arguments* (arguments are separate by `;`)
margin-top: 200px; 
margin-bottom: 200px; 
margin-left: 200px; 
margin-right: 200px; 
font-size: 3em; 
color: red;
background-color:
font-style: italic
border: 4px dotted red;
border-bottom: 4px dotted red;
letter-spacing: 2px;
opacity: 0.3;
padding-left: 50px;
position: (absolute, static,relative);
  right:5px; 
  top: 5px
  bottom:
  top:
  z-index:
  
  
  
##### CSS Examples

##### Section heads
.sectionhead {
  font-size: 1em;
  text-align: center;
  color: $presentation-heading-color;
  font-family: $presentation-heading-font;
  background-color: $body-bg;
  margin: 1px;
  padding: 2px 2px 2px 2px;
  width: 120px;
  border-left: 10px solid;
}



##### How to complete missing data


###### Set group_by() .drop to FALSE

If data is a factor and you have filtered out the data 
(but its listed as level), then you can set the `.drops` argument to `FALSE` in the `group_by()` and `summarize()` pattern and levels that have been filtered or do not have values will show up (by default these are droped `.drop=TRUE`)

```{r}


missing_data <- diamonds %>% 
  filter(cut!='Ideal')


missing_data %>% 
  group_by(cut) %>% 
  summarise(n=n())


```

-   Since `Ideal` has been filtered out we don't see it in the table above
-   Howver, sometimes we do want to show the values are zero


```{r}

missing_data %>% 
  group_by(cut,.drop=FALSE) %>% 
  summarize(n=n())


```
- Setting `.drop` to `FALSE` will now include the missing levels and their count (0) 

###### Use complete()


```{r}

missing_data <- mpg %>% 
  filter(!((manufacturer=='audi')&(year==2008)))

missing_data %>% 
  count(manufacturer,year)

```

- we see that the `audi` and `2008` are missing as we have filtered it out

- If we want to show these missing missing values, we can use complete() which will look ensure for every value in each column pair, there is entry

```{r}

#shows the missing values

missing_data %>% 
  count(manufacturer,year) %>% 
  complete(manufacturer,year)

# replaces the missing values with an entry
missing_data %>% 
  count(manufacturer,year) %>% 
  complete(manufacturer,year,fill = list(n=0))

roxygen2::
```

Original post on [Albert Rapp's Twitter](https://twitter.com/rappa753/status/1611754568123289601)



## How to create new packages

library(devtools)
library(usethis)


Create packge in the workspace that you want
devtools::create_package("package_name")

then do usethis::use_r("function_name") to create a script, put yoru function there

then do code>insert roxygen skeleton to comment code
load_all() to load code


## how to use nested map

- if access arguments in a column just access column
- if access nested data frame you can use .x$var
- if column is quoted then use .data[[var]] and define var="quoted var"
- if column is inserted as position, you can use .data[[pos]] for position arguments

*Considerations*
-   need to be mindful about the fomrula
      
    -   tidy friendly formula
    -   has data argument
    -   does not have data argument
-   helpful links
[r4epi purr](https://www.r4epi.com/using-the-purrr-package.html)
[paulvan purr](https://paulvanderlaken.com/tag/purrr/)
framework
[jenny bc](https://jennybc.github.io/purrr-tutorial/)
[got](http://zevross.com/blog/2019/06/11/the-power-of-three-purrr-poseful-iteration-in-r-with-map-pmap-and-imap/)
[presnetation on advance purr functions](https://hendrikvanb.gitlab.io/slides/purrr_beyond_map.html#8)

[going off the map](https://hookedondata.org/posts/2019-01-09_going-off-the-map-exploring-purrrs-other-functions/)
[scraping data with purr](https://colinfay.me/purrr-web-mining/)
[rowwise purr](https://thatdatatho.com/rowwise-purrr-pmap-apply-split-apply-combine/)

[modify_if purr](https://thatdatatho.com/rowwise-purrr-pmap-apply-split-apply-combine/)
[additional purr tricks](https://www.brodrigues.co/blog/2017-03-24-lesser_known_purrr/)
[overview presentatoin](https://shannonpileggi.github.io/iterating-well-with-purrr/#/title-slide)
[additional purr functoins](https://rstudio-pubs-static.s3.amazonaws.com/602410_2171106b3c7d429b96e606e8e41960a4.html)
[examples of purr](http://joshuamccrain.com/tutorials/purrr/purrr_introduction.html)
[applied purr](https://www.weirdfishes.blog/blog/practical-purrr/)
[addition purr fucntions](https://www.emilhvitfeldt.com/post/2018-01-08-purrr-tips-and-tricks/)

```{r}
library(tidyverse)

df <- mtcars %>% 
  rownames_to_column("brand")

nested_df <- df %>% 
  group_by(gear) %>% 
  nest()

nested_df


  



```

-   nested dataframe with names of columns directly in formula

```{r}
df

nested_df <- df %>% 
  group_by(gear) %>% 
  nest()

agg_fun <- function(df) {
  sum(df$hp - df$drat)
}

nested_df %>% 
  mutate(test=map_dbl(data,agg_fun))
  
  
```
-   nested dataframe with names passed on as arguments in formula (unquoted)

```{r}
nested_df <- df %>% 
  group_by(gear) %>% 
  nest()

agg_fun <- function(df,var1,var2) {
 x=df %>% pull({{var1}})
 y=df %>% pull({{var2}})
 print(sum(x - y))
}

nested_df %>% 
  mutate(test=map_dbl(.x=data,~agg_fun(df=.x,hp,drat)))

```



```{r}
library(tidyverse)
tibble(
  x = 1:10,
  y = 100:109,
  r = cor(x, y))


```


-   nested dataframe with with names reference from source data as arguments in formula

```{r}
nested_df <- df %>% 
  group_by(gear) %>% 
  nest()

agg_fun <- function(df,var1,var2) {
 x=df %>% pull({{var1}})
 y=df %>% pull({{var2}})
 print(sum(x - y))
}

sel_fun <- function(df,...) {
  df <- df %>% 
    select(any_of(c(...)))

sum(df[[1]] - df[[2]])
  
}



nested_df %>% 
  mutate(test=map_dbl(.x=data,
                      ~sel_fun(df=.x,"qsec","mpg")
                      )
         )

```

-   nested dataframe with names names references as list from source data as arguments in formula



-   trying to do aggregated function (eg return a single value) to each column of a table based on fixed value
  -   all_of() for quoted columns
  -   any_of() for quoted columns
  -   use this when you want to select the column names from vector
  
- quos or quos or quo_name()
    
    -   example
      map_dfr(
      .x = quos(age, ht_in, wt_lbs, bmi),
      .f = continuous_stats
      )
      
- if you use this then you pair this with {{}} - you don't need to use "" marks
      
[quos and purr](https://www.r4epi.com/using-the-purrr-package.html)


We haven’t seen the quos() function before. It’s another one of those tidy evaluation functions. You can type ?rlang::quos in your console to read more about it. When we can wrap a single column name with the quo() function, or a list of column names with the quos() function, we are telling R to look for them in the data frame being passed to a dplyr verb rather than looking for them as objects in the global environme
  

```{r}

#delcare variables
vars <- c("mpg","wt","magic")
pos <- c(2,7)


# direct and unquoted based on name or position
df %>% mutate(across(c(mpg,wt),
              max))
df %>% mutate(across(c(2,7),mean))
# direct and quoted
df %>% mutate(across(c("mpg","wt"),
              max))

# indirect and quoted based on name or postion
df %>% mutate(across(any_of(vars),mean))
df %>% mutate(across(all_of(pos),mean))




```

-   trying to do aggregated function (eg return a single value) to each column of a table based on patterns in the column name


```{r}

#delcare variables
vars <- c("mpg","wt","magic")
pos <- c(2,7)


# direct and unquoted based on name or position
df %>% mutate(across(starts_with("h"),
              max))

df %>% mutate(across(matches("^h"),
              max))

df %>% mutate(across(ends_with("t"),
              max))

df %>% mutate(across(matches("t$"),
              max))
```


-   trying to do multiple things to each column of a table base on column type

```{r}

df %>% mutate(across(starts_with("h"),
              max))



```

-   trying to do multiple things to each column of a table base on summarized criteria of a column

    - we can use predicate functions with  where() to identify columns that meet those criteria and select those columns

```{r}

mean_check_2 <- function(x,threshold) {
  threshold < mean(x,na.rm=TRUE)
}
select
mtcars %>% select(where(~mean_check_2(.,10)))
```


-   trying to do multiple things to each column of a table base on a single column attribute (Eg. missing data) 
  -   goal here is to come up with predicate function that tests the attribute you want and then apply that function to each column

```{r}

missing_df <- tibble(x=1:10,
                     y=c(1:5,NA_integer_,NA_integer_,1:3),
                     z=2:11,
                     a=3:12)

missing_fun <- function(x,threshold) {
  
  sum(is.na(x)) > threshold
  
}


missing_df %>% select(where(~!missing_fun(.,1)))

```

-   trying to do multiple things to each column of a table base on a multiple column attributes (eg. three consecutive rows)


```{r}

```

-   trying to do multiple things to each column of a table base on multiple column location attributes
-   trying to do multiple things on a pair(s) of columns of the above




```{r}

library(tidyverse)

# nested dataframe by geras
df <- mtcars %>% 
  group_by(gear) %>% 
  nest()


# custuom function
count_quoted <- function(data,var) {
  data %>% 
    count(.data[[var]])
}

# rowwise application
df %>% 
  rowwise() %>% 
  mutate(freq=list(
    count_quoted(data,"carb")
                   )
    ) %>% 
  pull(freq)

#map() application
df %>% 
  mutate(freq=map(data,~count_quoted(.x,"carb")
                   )
    ) %>% 
  pull(freq)

```

-   pe

-   
regex


"\\b" is the word boundary eg word1 and word2 will



if_any or if_all() for summarized attribute test against another test

```{r}
library(dplyr, warn.conflicts = FALSE)
library(palmerpenguins)


big <- function(x) {
  x > mean(x, na.rm = TRUE)
}

penguins %>% 
  filter(if_all(contains("bill"), big))

penguins %>% 
  filter(!is.na(bill_length_mm)) %>% 
  mutate(
    category = case_when(
      if_all(contains("bill"), big) ~ "both big", 
      if_any(contains("bill"), big) ~ "one big", 
      TRUE                          ~ "small"
    )) %>% 
  relocate(last_col())

```

## one note

-   purrr?
-   vscode?
-   timeseries?

```{r}
lass <- tibble(
  ht_in = c(70, 63, 62, 67, 67, 58, 64, 69, 65, 68, 63, 68, 69, 66, 67, 65, 
            64, 75, 67, 63, 60, 67, 64, 73, 62, 69, 67, 62, 68, 66, 66, 62, 
            64, 68, NA, 68, 70, 68, 68, 66, 71, 61, 62, 64, 64, 63, 67, 66, 
            69, 76, NA, 63, 64, 65, 65, 71, 66, 65, 65, 71, 64, 71, 60, 62, 
            61, 69, 66, NA),
  wt_lbs = c(216, 106, 145, 195, 143, 125, 138, 140, 158, 167, 145, 297, 146, 
             125, 111, 125, 130, 182, 170, 121, 98, 150, 132, 250, 137, 124, 
             186, 148, 134, 155, 122, 142, 110, 132, 188, 176, 188, 166, 136, 
             147, 178, 125, 102, 140, 139, 60, 147, 147, 141, 232, 186, 212, 
             110, 110, 115, 154, 140, 150, 130, NA, 171, 156, 92, 122, 102, 
             163, 141, NA)
)


cor.test(lass$ht_in,lass$wt_lbs,na.rm=TRUE) %>% broom::tidy()
```


### purrr framework

[purr gapminder example](https://www.rebeccabarter.com/blog/2019-08-19_purrr#simplest-usage-repeated-looping-with-map)


```{r}
gapminder_orig <- read.csv("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder-FiveYearData.csv")


gapminder_orig
```


[purr good pictures](https://dcl-prog.stanford.edu/purrr-extras.html)


## test



```{r}

eda_fun <- function(df) {
df %>% tibble(distinct=n_distinct(.x),
                   range=range(.x),
                   avg=mean(.x),
                   median=median(.x)
                   )
  )
}


  tibble(
    name=colnames(mtcars$mpg),
    class=class(mtcars$mpg),
    distinct=n_distinct(mtcars$mpg),
    range=paste0(range(mtcars$mpg)[1]," - ",range(mtcars$mpg)[2]),
    avg=mean(mtcars$mpg),
    median=median(mtcars$mpg),
    missing=sum(is.na(mtcars$mpg))
    )

  
gapminder_nested <- gapminder_orig %>% 
  group_by(continent) %>% 
  nest()  

args <- tibble(cols=c("pop","lifeExp","gdpPercap"))

new_df <- gapminder_nested %>%
  crossing(args)

cus_mean <- function(df,cols) {

  cols <- quos(cols)
  
  mean(df$cols)
  
}

new_df %>% 
mutate(avg=
           map2_dbl(.x=data,
                    .y=cols,
                    function(.x,.y) .x %>% pull(.y) %>% mean(.,na.rm=TRUE),.progress = TRUE
                    )
         )
```
## purr continueod


group_map()

this works more like summarize on a grouped dataframe, whereas regular map() works more lke mutate (in that you need a mutate to add a couple)

```{r}

library(dplyr)
library(purrr)

df <- data.frame(
  group = c("A", "A", "B", "B", "B"),
  value = c(1, 2, 3, 4, 5)
)

df %>%
  group_by(group) %>%
  group_map(~ mean(.$value))

df %>%
  split(.$group) %>%
  map(~ mean(.$value))


df %>%
  group_by(group) %>%
  nest() %>%
  mutate(mean = map_dbl(data, ~ mean(.$value))) %>%
  select(group, mean)




```
# time series & machine learning

## kmeans



- step 1: 
organize  data into normalied format

whatever we tryign to study (eg. product purchases) then you fill in the percent of that produc tby the dimenion you want (customer), eg customer by product

- step 2:
use kmeans function with table and initial group selection

-only can have numeric input (so create with explantory varibale then deselect variable (eg. customers))

kmeans_obj = kmeans(nomarlized_tbl,
centers=num__of_groups,
iter.max=something,
nstarts= higher is better (say 100))



kmeans_obj$center = all various centers

kmeans_obj$cluster=classify each obs to a clsuter

step 3: apply kmeans_obj to data

broom::augment(kmeans_obj,normalized_tbl)

step4: how to intrpret
tot.withinss is metric of total squared varition (lower is better), 

broom::glance(kmeans_obj)]
g

-   step5 optimze by setting up a grid

create function that calcules kmeans with cneters (and nstarts if you want) to be iterated with an input table and pipe this into glance

- plot tot.withinss and cneters to see which centers is optimized

# kmeans example

```{r}
devtools::load_all("/home/hagan/R/fpaR")
sales_by_store_product <- fpaR::contoso_fact_sales %>% 
  janitor::clean_names() %>% 
  mutate(sales_amount=unit_price*sales_quantity) %>% 
  group_by(store_key,product_key) %>% 
  summarise(sum_usd=sum(sales_amount),.groups = "drop") %>% 
  arrange(product_key) %>% 
  group_by(store_key) %>% 
  mutate(prop_usd=sum_usd/sum(sum_usd)) %>% 
  select(store_key,product_key,prop_usd) %>% 
  pivot_wider(names_from=product_key,values_from=prop_usd,values_fill = 0,names_repair = janitor::make_clean_names)


kmeans_obj <- sales_by_store_product %>% 
  select(-store_key) %>% 
  kmeans(centers=15,nstart=100,iter.max = 100)

stores_with_kmeans <- kmeans_obj %>% broom::augment(data=sales_by_store_product) 
kmeans_obj %>% broom::glance()


kmeans_mapper <- function(centers=3) {
  sales_by_store_product %>% kmeans(x=.,centers=centers,nstart = 100,iter.max=100)
  
}

grid_kmeans <- tibble(centers=1:15) %>% 
  mutate(kmeans_obj=map(.x=centers,~kmeans_mapper(.x)),
         kmeans_output=map_df(kmeans_obj,broom::glance)
  )

grid_kmeans %>%
  unnest(kmeans_output) %>% 
  select(centers,tot.withinss) %>% 
  ggplot(aes(x=centers,y=tot.withinss))+
  geom_point()+
  geom_line()

```

## umap

very simliar to kmeans just without addition arguments
stregth is merge the kmeans output with the kmeans output so that you can mpa them together
then manually inspect original data frame by the new clusters to see what patterns or attributes (eg. can sort by some cumsum and attributes to look for patterns)


step1 
- use same normalized model
- only select numeric data
- pass through to umeans

step2
- take hte umeans$layout covert to tibble, rename columns and bind with framing column




```{r}

library(umap)
umap_obj <- sales_by_store_product %>% 
  select(-store_key) %>% 
  umap()

umap_tbl <- umap_obj$layout %>%
  as_tibble() %>% 
  set_names(c("x","y")) %>% 
  bind_cols(sales_by_store_product) %>% 
  select(store_key,x,y)

umap_tbl %>% 
  ggplot(aes(x,y))+
  geom_point()
  
kmeans_result_tbl <-   grid_kmeans %>%
    filter(centers==8) %>% 
    pull(kmeans_obj) %>% 
    pluck(1) %>% broom::augment(sales_by_store_product) %>% 
    relocate(last_col()) %>% 
    select(.cluster,store_key)


kmeans_result_tbl %>% 
  left_join(umap_tbl,by="store_key") %>% 
  ggplot(aes(x=x,y=y,col=.cluster))+
  geom_point()

```


## combine umeans and kmeans

