# Chapter 1: Let's get started

# Named Objects

::: callout-note
in this package, we will use the following libraries

library(tidyverse)
:::

```{r}
#|echo: false

pacman::p_load(tidyverse)

```


### how to create and reference objects and generate series

Although not the most glamorous place to start, learning how to create named objects, reference and manipulate them as well generate data becomes incredibly useful down the road when combined with other techniques and workflows.

in R this is main way you save, reference, and manipulate your data

you use the `<-` (ctrl + - ) or the `=` symbol to assign a result to an object

```{r}


#| echo: true
object <- 1+1
```

Whatever name you type will become the object's name, if you type in the object, it will return whatever you assigned to the name

Although this is fairly intuitive to understand, if you are coming from an Excel background, it will take time to develop this habit of assigning names to results because in Excel, you can refer to a result by its cell location by physically pointing a formula to a input.

For quick and simple calculations, this will be quicker but if you try to do something even beyond a simple calculation, you will see it becomes very difficult to manage, easy to make mistakes and reduces your productivity as you have to constantly point to those input cells or have to remember non-intuitive names `eg C5` .

### In Excel
In excel, you can also name a cell by putting your cursor over the cell and click on the named box in the top left.

This is a useful technique for commonly referenced cells but the "workflow" is a bit nonstandard. You can use F5 to quickly name an object

To keep track of all the names, you can press F5 and you will see a dialog box that will help you manage all of your names.

When you want to refer to that cell, you can simply type in the cell name.

although useful for intermediate analysis, I would say from a workflow perspective, it isn't "easy" as you have to constantly break your workflow but it will make your work more readable.

### Types of data

int
character
date
NA
NULL


```{r}
character_example <- "test"
numeric_example <- 1
date_example <- as.Date("2012/2/12")
integer_example <- 2L


```


you don't pay attention to data types in Excel as much until you have problems -- it is a strength and weakness. You can throw any values together and excel doesn't care (until you try to do something with it).

R (and most other languages) really care about data types -- in fact they won't make you do anything unless data types are aligned



## structure

who actually knows but the best explanation is that this stands for concatenate -- which basically means brings things together.

This is a new concept for excel users --, we can just type things in a cell and it works. 

In R -- you can't do that, we need to add some structure to our data in the form either a vector or list or data frame (which is just a collection of lists)

This may feel overwhelming -- in reality it will feel very intuitive -- only later on will some nuance really distinguish between these types

### Vector

```{r}
a <- vector()

a <- 1:10

a

b <- c(1:10)
?vector()
identical(a,b)
```


"logical"
,"integer"
, "double"
, "complex"
, "character"
, "raw"
,"list"
, "NULL"

```{r}
a <- c(1:10)
b <- rlang::list2(1:10)

a==b[[1]]
identical(a,b[[1]])
```

lists can be blessing and course

```{r}
li
```


## classes of data

Checking clases of data isn't a routine action but if you have errors, its helpful in solving this

we can use `class()` to check the class of object are you can be specifc with the is._ type funcitons

is.numeric()
is.character()
is.Date()



```{r}
class(character_example)
class(numeric_example)
class(date_example)
class(integer_example)

base::is.character()
base::is.integer()
base::is.numeric()
lubridate::is.Date(date_example)

typeof(character_example)

```


### Excel



## Type conversions --


So what happens if you combine a character with interger?

You get what is called a type conversin -- eg. R will automatically convert one character type to another

```{r}


c(1:10,"a")
```



### how to generate data

## In R

This may seem useful but I assure its very useful

list of numbers

```{r}
1:10

10:1
```

alternatively you can use the function `seq()`


```{r}
seq(1,100,10)

seq_along(1:100)

seq_len(10)

sequence(c(5, 5), from=1L, by=2L)

seq.int(1,10,1)


seq.Date(
  from=as.Date("2021-01-01")
  ,to = as.Date("2021-01-03")
  ,by="days"
)


sample(1:100,size=30,replace = TRUE)
## rep

rep_named(words[1:100],1:100) %>% stack()

words


## runif
n=100
dunif(.2, min = 0, max = 1, log = FALSE)
punif(q, min = 0, max = 1, lower.tail = TRUE, log.p = FALSE)
qunif(p, min = 0, max = 1, lower.tail = TRUE, log.p = FALSE)
runif(n, min = 0, max = 1)
## distributions

#build in dataaets
letters

LETTERS

state.abb


## sample

test <- rlang::list2(a=1:10
             ,b=1
             ,c=1:4)

test$c

test[[1]] - test[[3]]

test[["c"]] - test[["a"]]



```

###

```{r}
library(rlang)
a<- list2()

a <- list(b=1:10,c=1)

a <- list(c=2:10)

append(a,values=list2(a=1:2,z=1:10),after = 2)
```

when we create lists or vectors, sometimes you want to add to it- in reality not really but we want to get you more comfortable with lists and vectors and introduce some simple functions that can be helpful later in your journey

`append()` is a great function that adds to an existing function

append(
vector or list
,new entry
,after= number of list
)

In practice you won't use this much in your beginning analysis -- it can be helpful later on when you start to write functions but for now its just to get you use to the command line and scripting and take away some of the angst

## subsetting

how do we extract from vectors

well this is where these bracket things come from basically you are can look inside and either call out by number or name the object you want

```{r}

list2(
  a=1:10
  ,b=words[10]
  ,c=words[1:3]
  ,d=10:1
  )

```


## recyling


```{r}
data()
freeny

map_chr(rep(1,100),function(.x) sample(words,.x,replace = TRUE))
```


```{r}

data()
luv_colours %>% 
  ggplot() +
  geom_point(aes(x=u,y=v,col=col))+
  scale_color_identity()+
  theme_classic()


table2
```

## practice!

stimulate a dice roll?

```{r}
sample(1:6,1,replace=TRUE)

```


## Classes of data




### how to create and named Tables

##### In R

regardless if the data object is a cell or a table (or actually any type of object) the naming methodology is the same.

however creating a table isn't as straight forward because you don't have the same GUI.

data entry is not R (or any scripting languages) strong suit - which doesn't mean you can't do it but if you are strictly entering in data (of great length), use excel.

however, below shows you how to create tables

#### tibble / data frame

```{r}
library(tidyverse)
df_example <- tibble(
  x=1:10
  ,y=2:11
  ,z=3:12
  )

```

#### Tribble

When trying to do data entry that you need to reference columns wise & rowwise, this can challenging, so thats why there is a `tribble`, a handy way to still see your columnwise and rowwise data together

::: callout-note
there is a concept of vectorization, which for now, we will define as recycling. Which is to say if I put one column with 10 elements, and another column with only 1 element, tibble will recycle the "1" until it length of the tibble.
:::

```{r}
df_tribble <- tribble(
  ~col1,~col2,~col3,
  1,2,3,
  5,6,7,
  8,9,10
  
  
)
```

another methodology which can be useful are the `data.edit` and

```{r}

```

::: callout-tip
## Naming Convention
what naming convention should we use in R? This is where you will start to see naming conventions that you are not familiar with such as snake_case, kebab-case, or Upper_camel.

for this book, we will use snake_case (but you use whatever is easist for you).

coming up with names for variables will be challenging as you will either type too little or too much, over time you will get your own convention. It is not important what is optimal, more important to focus on convention (means consistency) so that when you read something you immediately know what it is.

for this book if its a regular R object we will append the object with `_tbl` to help keep track of objects
:::

### Excel Skills to be covered

In excel, this is for more straight forward -- you simply put the data into cells. Highlight the area and then click insert>table

From there, Its a good practice to name the table (this will make referencing the table much easier and straight forward).

Again for quick and simple analysis and one time analysis -- not necessary



## adding columns to Tables

### In R

In R we can add to our base table by using the dplyr function `mutate()` which basically takes two arguments, the name of the column and the content of the column.

```{r}
tibble(
  x=1:10
  ,y=letters[1:10]
) %>% 
  mutate(z=1)
```


Probably the most confusing part of this for you will be the name mutate -- it doesn't sound rational or logical but in programming languages there is a concept called mutate where you change an object. 

That may or not help you in remembering how to change an object but you will absolutely be using this function alot

Secondly, this is your first introduction to the pipe! It looks funny but soon you will be super addicted to it.

I said there are two key parameters for `mutate()` but actually, there is three mutate(data,column_name,value). However when I used it, I didn't specificy the data object. 

That is because the %>% puts the left hand side as the first argument in the right handside function.

so `mutate(tbl,z=1:10)` because `tbl %>% mutate(z=1:10)`. Additionally you can do `tbl |> mutate(.,z=1:10)`.

Ok -- now for a first point of confusion -- you will sometimes see the pipeline written as `|>` and sometimes `%>%` 

::: callout-tip
## history


:::



## Tidyselect verbs -- first superpower

```{r}

```


## Adding rows to a table

### In R 

You can do the following

-   Make a table of values and bind it with `bind_rows()`
-
-   

```{r}

dplyr::rows_append(),dplyr::rows_insert(),dplyr::rows_update()



```

### In Excel

Simply add the values you want at the bottom of the table


## Referencing named objects


### In R

use the `ls()` funciton to see all the objects you named

in Excel you use F5


index() 



tab references 

how to edit named objects 

data Entry Form add in 

existing data 

new series of Numbers 

new series of Dates 

flash Fill tips 

adding/subtracting/dividing/multiplying columns 


```{r}
diamonds |> 
  mutate(ratio=table/depth)
```

one formula works against all rows in a column 

rowwise equvalant




colwise equvailant 

mean 

median 

sum 

count 

aggregate 

error handling 

iferror() 

nA vs. null vs. ref 

array \[what did I mean by this\]

## R skills to be covered

tibbles 

lists 

vectors 

dataeditR 

arrange() 

generating data 

-numbers 

-letters 

-dates 

-rep() 

-seq() 

-mutate

::: {#special .sidebar}
::: warning
here is a warning.
:::

more content.
:::



<<<<<<< HEAD
#tidymodels

split data & sample folds

1.  rsample::initial_sample(dataframe,strata=variable)

2.  rsample::training(split_data)

3.  rsample::testing(split_data)

4.  vfold_cv(df_testing, strata = type)

#### recipe to create formula and transform daata

1.  recipes::recpe(outcome \~ predictor,data=train_df)

#### select the ml model and type

parsnip::logistic_reg() %\>% set_mode() %\>% set_engine()

### workflow

workflow() %\>% add_recipe(rec) %\>% add_model(mod)




i(rgb) will take the literal names f the columns and apply the colors https://www.youtube.com/watch?v=CTWkJrvfRBc minute 37:42

#### Metadata for columns

resource:

\[link to original source\](https://www.pipinghotdata.com/posts/2022-09-13-the-case-for-variable-labels-in-r/)

    penguins_metadata <- tribble(
        ~variable,           ~variable_label, 
        "species",           "Penguin species",
        "island",            "Island in Palmer Archipelago, Antarctica",
        "bill_length_mm",    "Bill length (mm)",
        "bill_depth_mm",     "Bill depth (mm)",
        "flipper_length_mm", "Flipper length (mm)",
        "body_mass_g",       "Body mass (g)",
        "sex",               "Penguin sex",
        "year",              "Study year"
    )

`labelled::generate_dictionary()`

generates a data dictionary

subsittute labels for points in graph

to substitute variable labels for variable names on your input variables, add [`ggeasy::easy_labs()`](https://jonocarroll.github.io/ggeasy/reference/easy_labs.html) to your ggplot.

    penguins_labelled <- penguins |> 
      set_variable_labels(
        species           = "Penguin species",
        island            = "Island in Palmer Archipelago, Antarctica",
        bill_length_mm    = "Bill length (mm)",
        bill_depth_mm     = "Bill depth (mm)",
        flipper_length_mm = "Flipper length (mm)",
        body_mass_g       = "Body mass (g)",
        sex               = "Penguin sex",
        year              = "Study year"
      )

    penguins_labelled <- penguins |> 
      set_variable_labels(!!!penguins_labels)



#### Quarto style guides

as best I can tell there are 3 ways to format presentations


`:::{.args input1="..." input2="...}`

impact text
`:::`
- this appears to reference built in quarto arguments
- they start with .args followed by inputs into that args

{style="args:..."}`
- this appears to reference CSS style arguments

##### Change changing style within string
if you want to edit a single word or string of  words **within a string** 


[existing text]{style="args:..."}


or if you want to edit a whole string (not just parts of it)

`:::{style="args:..."}` 
insert text
`:::` 

*popular arguments* (arguments are separate by `;`)
margin-top: 200px; 
margin-bottom: 200px; 
margin-left: 200px; 
margin-right: 200px; 
font-size: 3em; 
color: red;
background-color:
font-style: italic
border: 4px dotted red;
border-bottom: 4px dotted red;
letter-spacing: 2px;
opacity: 0.3;
padding-left: 50px;
position: (absolute, static,relative);
  right:5px; 
  top: 5px
  bottom:
  top:
  z-index:
  
  
  
##### CSS Examples

##### Section heads
.sectionhead {
  font-size: 1em;
text-align: center;
color: $presentation-heading-color;
font-family: $presentation-heading-font;
background-color: $body-bg;
margin: 1px;
padding: 2px 2px 2px 2px;
width: 120px;
border-left: 10px solid;
}   






[1] "hello"

a asfds
a dog and elephant
a list of things
a safdfs
aas sdf
b dsfsfa
cadfaa
d adfas


```{r}



```




##### How to complete missing data


###### Set group_by() .drop to FALSE

  If  data is a factor and you have filtered out the data 
(but its listed as level), then you can set the `.drops` argument to `FALSE` in the `group_by()` and `summarize()` pattern and levels that have been filtered or do not have values will show up (by default these are droped `.drop=TRUE`)

```{r}


missing_data <- diamonds %>% 
  filter(cut!='Ideal')


missing_data %>% 
  group_by(cut) %>% 
  summarise(n=n())


```

-   Since `Ideal` has been filtered out we don't see it in the table above
-   Howver, sometimes we do want to show the values are zero


```{r}

missing_data %>% 
  group_by(cut,.drop=FALSE) %>% 
  summarize(n=n())


```
- Setting `.drop` to `FALSE` will now include the missing levels and their count (0) 

###### Use complete()


```{r}

missing_data <- mpg %>% 
  filter(!((manufacturer=='audi')&(year==2008)))

missing_data %>% 
  count(manufacturer,year)

```

- we see that the `audi` and `2008` are missing as we have filtered it out

- If we want to show these missing missing values, we can use complete() which will look ensure for every value in each column pair, there is entry

```{r}

#shows the missing values

missing_data %>% 
  count(manufacturer,year) %>% 
  complete(manufacturer,year)

# replaces the missing values with an entry
missing_data %>% 
  count(manufacturer,year) %>% 
  complete(manufacturer,year,fill = list(n=0))

```

original post on [Albert Rapp's Twitter](https://twitter.com/rappa753/status/1611754568123289601)



orkflow

workflow() %\>% add_recipe(rec) %\>% add_model(mod)


=======
`tibble::enframe()` takes a named list and turns into a tibble
>>>>>>> a548534aa2ba9c3ce865bab0e837a2867fc5c783


i(rgb) will take the literal names f the columns and apply the colors https://www.youtube.com/watch?v=CTWkJrvfRBc minute 37:42

#### Metadata for columns

resource:

\[link to original source\](https://www.pipinghotdata.com/posts/2022-09-13-the-case-for-variable-labels-in-r/)

    penguins_metadata <- tribble(
        ~variable,           ~variable_label, 
        "species",           "Penguin species",
        "island",            "Island in Palmer Archipelago, Antarctica",
        "bill_length_mm",    "Bill length (mm)",
        "bill_depth_mm",     "Bill depth (mm)",
        "flipper_length_mm", "Flipper length (mm)",
        "body_mass_g",       "Body mass (g)",
        "sex",               "Penguin sex",
        "year",              "Study year"
    )

`labelled::generate_dictionary()`

generates a data dictionary

subsittute labels for points in graph

to substitute variable labels for variable names on your input variables, add [`ggeasy::easy_labs()`](https://jonocarroll.github.io/ggeasy/reference/easy_labs.html) to your ggplot.

    penguins_labelled <- penguins |> 
      set_variable_labels(
        species           = "Penguin species",
        island            = "Island in Palmer Archipelago, Antarctica",
        bill_length_mm    = "Bill length (mm)",
        bill_depth_mm     = "Bill depth (mm)",
        flipper_length_mm = "Flipper length (mm)",
        body_mass_g       = "Body mass (g)",
        sex               = "Penguin sex",
        year              = "Study year"
      )

    penguins_labelled <- penguins |> 
      set_variable_labels(!!!penguins_labels)



#### Quarto style guides

as best I can tell there are 3 ways to format presentations


`:::{.args input1="..." input2="...}`

impact text
`:::`
- this appears to reference built in quarto arguments
- they start with .args followed by inputs into that args

{style="args:..."}`
- this appears to reference CSS style arguments

##### Change changing style within string
if you want to edit a single word or string of  words **within a string** 


[existing text]{style="args:..."}


or if you want to edit a whole string (not just parts of it)

`:::{style="args:..."}` 
insert text
`:::` 

*popular arguments* (arguments are separate by `;`)
margin-top: 200px; 
margin-bottom: 200px; 
margin-left: 200px; 
margin-right: 200px; 
font-size: 3em; 
color: red;
background-color:
font-style: italic
border: 4px dotted red;
border-bottom: 4px dotted red;
letter-spacing: 2px;
opacity: 0.3;
padding-left: 50px;
position: (absolute, static,relative);
  right:5px; 
  top: 5px
  bottom:
  top:
  z-index:
  
  
  
##### CSS Examples

##### Section heads
.sectionhead {
  font-size: 1em;
  text-align: center;
  color: $presentation-heading-color;
  font-family: $presentation-heading-font;
  background-color: $body-bg;
  margin: 1px;
  padding: 2px 2px 2px 2px;
  width: 120px;
  border-left: 10px solid;
}



##### How to complete missing data


###### Set group_by() .drop to FALSE

if data is a factor and you have filtered out the data 
(but its listed as level), then you can set the `.drops` argument to `FALSE` in the `group_by()` and `summarize()` pattern and levels that have been filtered or do not have values will show up (by default these are droped `.drop=TRUE`)

```{r}


missing_data <- diamonds %>% 
  filter(cut!='Ideal')


missing_data %>% 
  group_by(cut) %>% 
  summarise(n=n())


```

-   Since `Ideal` has been filtered out we don't see it in the table above
-   Howver, sometimes we do want to show the values are zero


```{r}

missing_data %>% 
  group_by(cut,.drop=FALSE) %>% 
  summarize(n=n())


```
- Setting `.drop` to `FALSE` will now include the missing levels and their count (0) 

###### Use complete()


```{r}

missing_data <- mpg %>% 
  filter(!((manufacturer=='audi')&(year==2008)))

missing_data %>% 
  count(manufacturer,year)

```

- we see that the `audi` and `2008` are missing as we have filtered it out

- If we want to show these missing missing values, we can use complete() which will look ensure for every value in each column pair, there is entry

```{r}

#shows the missing values

missing_data %>% 
  count(manufacturer,year) %>% 
  complete(manufacturer,year)

# replaces the missing values with an entry
missing_data %>% 
  count(manufacturer,year) %>% 
  complete(manufacturer,year,fill = list(n=0))

```

original post on [Albert Rapp's Twitter](https://twitter.com/rappa753/status/1611754568123289601)



## How to create new packages

library(devtools)
library(usethis)


create packge in the workspace that you want
devtools::create_package("package_name")

then do usethis::use_r("function_name") to create a script, put yoru function there

then do code>insert roxygen skeleton to comment code
load_all() to load code
# time series & machine learning

## kmeans



- step 1: 
organize  data into normalied format

whatever we tryign to study (eg. product purchases) then you fill in the percent of that produc tby the dimenion you want (customer), eg customer by product

- step 2:
use kmeans function with table and initial group selection

-only can have numeric input (so create with explantory varibale then deselect variable (eg. customers))

kmeans_obj = kmeans(nomarlized_tbl,
centers=num__of_groups,
iter.max=something,
nstarts= higher is better (say 100))



kmeans_obj$center = all various centers

kmeans_obj$cluster=classify each obs to a clsuter

step 3: apply kmeans_obj to data

broom::augment(kmeans_obj,normalized_tbl)

step4: how to intrpret
tot.withinss is metric of total squared varition (lower is better), 

broom::glance(kmeans_obj)]
g

-   step5 optimze by setting up a grid

create function that calcules kmeans with cneters (and nstarts if you want) to be iterated with an input table and pipe this into glance

- plot tot.withinss and cneters to see which centers is optimized

# kmeans example

```{r}
devtools::load_all("/home/hagan/R/fpaR")
sales_by_store_product <- fpaR::contoso_fact_sales %>% 
  janitor::clean_names() %>% 
  mutate(sales_amount=unit_price*sales_quantity) %>% 
  group_by(store_key,product_key) %>% 
  summarise(sum_usd=sum(sales_amount),.groups = "drop") %>% 
  arrange(product_key) %>% 
  group_by(store_key) %>% 
  mutate(prop_usd=sum_usd/sum(sum_usd)) %>% 
  select(store_key,product_key,prop_usd) %>% 
  pivot_wider(names_from=product_key,values_from=prop_usd,values_fill = 0,names_repair = janitor::make_clean_names)


kmeans_obj <- sales_by_store_product %>% 
  select(-store_key) %>% 
  kmeans(centers=15,nstart=100,iter.max = 100)

stores_with_kmeans <- kmeans_obj %>% broom::augment(data=sales_by_store_product) 
kmeans_obj %>% broom::glance()


kmeans_mapper <- function(centers=3) {
  sales_by_store_product %>% kmeans(x=.,centers=centers,nstart = 100,iter.max=100)
  
}

grid_kmeans <- tibble(centers=1:15) %>% 
  mutate(kmeans_obj=map(.x=centers,~kmeans_mapper(.x)),
         kmeans_output=map_df(kmeans_obj,broom::glance)
  )

grid_kmeans %>%
  unnest(kmeans_output) %>% 
  select(centers,tot.withinss) %>% 
  ggplot(aes(x=centers,y=tot.withinss))+
  geom_point()+
  geom_line()

```

## umap

very simliar to kmeans just without addition arguments
stregth is merge the kmeans output with the kmeans output so that you can mpa them together
then manually inspect original data frame by the new clusters to see what patterns or attributes (eg. can sort by some cumsum and attributes to look for patterns)


step1 
- use same normalized model
- only select numeric data
- pass through to umeans

step2
- take hte umeans$layout covert to tibble, rename columns and bind with framing column




```{r}

library(umap)
umap_obj <- sales_by_store_product %>% 
  select(-store_key) %>% 
  umap()

umap_tbl <- umap_obj$layout %>%
  as_tibble() %>% 
  set_names(c("x","y")) %>% 
  bind_cols(sales_by_store_product) %>% 
  select(store_key,x,y)

umap_tbl %>% 
  ggplot(aes(x,y))+
  geom_point()
  
kmeans_result_tbl <-   grid_kmeans %>%
    filter(centers==8) %>% 
    pull(kmeans_obj) %>% 
    pluck(1) %>% broom::augment(sales_by_store_product) %>% 
    relocate(last_col()) %>% 
    select(.cluster,store_key)


kmeans_result_tbl %>% 
  left_join(umap_tbl,by="store_key") %>% 
  ggplot(aes(x=x,y=y,col=.cluster))+
  geom_point()

```


## assert that
assertthat packageschecks if column names are in a dataframe


## combine umeans and kmeans



### data masking

sym() is used to take a quoted input "x" and use it in a context without quotes. Need !!sym("x") to be used

or  rlang::englue("var) within the .data[[var]] context


.data[[rlang::englue("var")]]

.env$var to refer to grid table

```{r}
library(tidyverse)


col_numbers <- 1:ncol(diamonds)

  map(
    .x=col_numbers
      ,.f=~diamonds[.x:10]
  )

  split(diamonds, ~ cut) %>% 
    map(
      .x=.
      ,~.x %>% pull(price) %>% sum) %>% 
    stack()

 
    map(
      .x=col_numbers
      ,.f = ~mtcars[.x] %>% table()
      )
```