---
title: "time series"
---

```{r}
library(tidyverse)
library(timetk)
library(modeltime)
library(tidymodels)
pkgload::load_all("/home/hagan/R/fpaR")

df <- fpaR::contoso_fact_sales %>% 
  janitor::clean_names() %>% 
  mutate(
    date_key=mdy(date_key)
    ,rev=sales_quantity*unit_price
    )
```
# Simplified framework

1. summarize by date column
2. add dimensions that you want in learning
3. check and add correlations
4. feature engineer such as scaling and normalizing data, difference and logging log
5. add time related signatures
6. create test and train data sets for augmentation and back testing too 
7. select models
8. tune models
9. create model workflows
10. fit models


# how to create aggreation fucntion


-   `summarize_by_time()` works similar to group_by()+summarize

timetk::summarize_by_time(
.date_var=date column
.by= aggregation level such as day, month, 8 weeks, quarter, year
col_name=aggregation step (n())
across()
,.type=c("ceiling","round","floor")
)


```{r}
df_daily <- df %>%
  summarise_by_time(
    ,.date_var = date_key
    ,.by = "day"
    ,rev=sum(rev)
    ,vol=sum(sales_quantity)
    ,mean_up=mean(unit_price)
    ) %>% 
  mutate(price_realization=rev/vol)

```


```{r}

test_fun <- function(df,.date_var,.value) {
  df %>% 
    summarise_by_time(.,.date_var={{.date_var}},rev=sum({{.value}}),.by="day")
}


test_fun(df=df,.date_var = date_key,.value=rev)

```


##  how to pad time when there gaps in data series
- `pad_by_time()`  use to fill gaps with values in your time series

pad_by_time(
.date_var=data column
.by="day"
.pad_value=0
)



## plot times series
-   `plot_time_series()` used to plot times series in either plotly or ggplot (set interactive to FALSE for ggplot2)
-   You can use in combination with group_by() to facet or you can use the facet_vars argument
-   contains additional formatting and plotly arguments
-   date_var and value are the most critical 

group_by() %>% ## how to facet
plot_time_series(
.data_var
,value
,.color_bar=group_name # puts groups in color
,.facet_Vars=name ## also how to facet
,.smooth=TRUE (adds in polynominal regression)
,.smooth_period="30 days"
,.smooth_span= prop of data
,.interactive=FALSE # to create ggplot object
)

```{r}
df_daily %>% 
  plot_time_series(
    .date_var = date_key
    ,.value=rev
    ,.line_color = "red"
    ,.line_type = "dotted"
    ,.smooth_period = 30
    ,.smooth_color = "green"
    ,.plotly_slider = FALSE
    ,.color_var = "black"
      ,.x_lab = "test"
    ,.y_lab = "test2"
    ,.color_lab = "blue"
      ,.smooth_alpha = .3
    ,.title = "Hello"
    )
```

## Filter time series 
-   Use `filter_by_time()` to filter times series with greater flexibility 
-   you can put a start date with one granularity (2013) and end date with a different (2015-02-01) or use "start" and "end" for the min and max of the dataset



filter_by_time(
.date_var=data column
start_date= "2013" "2013-01" "2013-01-5" or mix and match
end_date="end"  or date
)

```{r}

df_daily %>% 
  filter_by_time(
    .date_var=date_key
    ,.start_date = "2007-06"
    ,.end_date ='end'
  )
```

## split into training and test set

-   Similar to tidymodels `split()` this will split a time series 
-   you assign the split data to an object and use it with tidymodels `train()` and `test()`
-   need to set cumulative true if you want the training set to be historical to the assessment period

time_series_split(
 date_var=date columns
 ,assess= "8 weeks" or 200
 ,initial = 150 # data points
 cumulative = TRUE  # sets the previous as training, false means you need to select training set
)

```{r}
df_split <- df_daily %>% 
  time_series_split(
    date_var = date_key
    ,assess = 200
    ,initial = 150
    ,point_forecast = FALSE
    ,cumulative = TRUE
  )



```

- you can use the `training()` or `testing()` on the split object to get the various split elements

-   To view of plot of the split object you first need to pass the split object to `tk_time_series_cv_plan()`  and then use `plot_time_Series_cv_plan()` 

```{r}

df_split %>% 
  tk_time_series_cv_plan() %>% 
  plot_time_series_cv_plan(
    .date_var = date_key
    ,.value = rev
    ,.interactive=FALSE
    ,.smooth = TRUE
    ,.title="Trainig vs. Testing set"
  ) +
  scale_y_continuous(labels = label_comma())

```
## create different time series models

-   Linear model 

linear_reg() 

-   Facbeook's prophet model
prophet_reg(
) %>% 
set_engine("prophet")


```{r}
lm_mod <- linear_reg()

prophet_mod <- modeltime::prophet_reg() %>% 
  set_engine("prophet")
```


how to use?

- fit models with a recipe (formula + featuere engineering) and fit to to data set (training or testing)

prophet_reg %>% 
set_engine("prophet") %>% 
fit(predition_Value ~ data_col
,data=training(splits))


```{r}
lm_fit <- lm_mod %>% 
  fit(rev~month(date_key)+vol*mean_up,data=training(df_split))

prophet_fit <-  prophet_mod %>% 
  fit(rev~date_key+vol,data=training(df_split))

```



## model time worfklow
creates table to hold various fitted models
model_time_table(
fitted_model_1
)

- applies fitted models against testing data
- .calabration_data has the results
- savve as calibration data
- Appears to work simliar to worflowsets but without the preprocessing just holds the modesl

```{r}
model_tbl <- modeltime_table(
  lm_fit,
  prophet_fit
)



```




model_tbl %>% 
modeltime_calibrate(
new_data=testing(splits)

)

```{r}
calib_tbl <- model_tbl %>% 
  modeltime_calibrate(new_data=training(df_split))
```


to see this in practice
calibration_tbl %>% 
modeltime_forecat(actual_Date=existing_tlb) %>% 
plot_modeltime_forecast()


```{r}
calib_tbl %>% 
  modeltime_forecast(new_data=testing(df_split),
                     actual_data = df_daily,conf_interval = 0.1) %>% 
  plot_modeltime_forecast()
```




## how to see forecasting accuracy

calibration_tbl %>% modeltime_accuracy

```{r}
calib_tbl %>% 
  modeltime_accuracy()
```



plot_seasonal_diagnotistic


## recipe

- you provide the transformation steps here
-   be mindful that your training and and testing datasets need the to be converted with the training steps so that when you back test or fits everythign works

    -   same transformation columns
    -   ability to transform predictions back to untransformed output
    -   take note of the transformation vectors
    -   you apply the steps to the dataset with prep %>% juice

recipe(formula,training(splits)) %>% 
step_timeseries_Signature() %>%
step_normalize(index.num,_year) %>% 
step_dummy(all_ordinal)
prep() %>% 
%>% juice (applies to the training step)


## create model element

linear_reg()

## create a workflow object
save as worflow

workflow %>% 
add_model(model_spec) %>% 
add_reciept(reciepe-sepc) %>% 
fit(training(splits)) 

calibration obj
modeltime_table() %>% 
modeltime_Calibration(testing(splits))


calibration_obj %>% modeltime_accuracy

calibration_obj %>% 
modeltime_forecast(
new_Data=testing(splits)
actual_Data=origional_ttbl) %>% 
plot_modeltime_forecast()



)

# overview of functions

## diagnostics functions

### ACF

-   helpful to detect correlations between lagged and how many lags to use in a model

plot_acf_diagnostics(
.lags="1 year", 25:100,100
.ccf_vars=columns:columns   
.show_ccf_vars-only=TRUE # only show cross correlations var
,.facet_ncol=number of column
)

```{r}
df_daily %>% 
  plot_acf_diagnostics(
    .date_var=date_key
    ,.value=rev
    ,.lags = 1:30
  )
```

## sesaonality

-   Helpful to see if there are certain seasonality drivers (eg. weekly, monthly, day of month, etc) effect

plot_seasonal_diagnostics(
feature_set=c(hour,weekday)
.geom="violin"
)


```{r}
df_daily %>% 
  plot_seasonal_diagnostics(
    .date_var=date_key
    ,.value = rev
      )

```


### anolomy

-   helpful to detect and plot anomalies detection

plot_anomaly_diagnostics(
.date_var=date_col
.value=value_col
.alpha= .01 #change anomaly sensitivity
.max_anolamoies =.01 # peak anomaly

)


```{r}
df_daily %>% 
  plot_anomaly_diagnostics(
    .date_var = date_key
    ,.value=rev
    ,.alpha=.2
  )
```


-   Get underlying anomaly data
tk::anomoly_diaogasitc(
.data_var=data_col
.alpha= .01 #change anomoly sensitivity
.max_anolomoies =.01 # peak anomoly
)


```{r}
df_daily %>% 
  timetk::tk_anomaly_diagnostics(
    .date_var = date_key
    ,.value=rev
    ,.alpha = .3
  )
```


## seasons decomposition

-   Useful to understand how anomoliges are being detcted
plot_stl_diagostics(


)

```{r}
df_daily %>% 
  plot_stl_diagnostics(
    .date_var=date_key
    ,.value=rev
    ,.feature_set = c("season","trend")
  )
```



## linear regression plot

-   shortcut function of `lm()` to help identify which features are meaningful or not

plot_time_series_regression(
.date_var=data column
.formula= value~ as.numeric(date_col)+
wday(date_col)+
month(date_col)
show_summary= TRUE, shows summary ouptut
)



```{r}
df_daily %>% 
  plot_time_series_regression(
    .date_var = date_key
    ,.formula = rev ~ vol+week(date_key)+mean_up+timetk::lag_vec(date_key,lag=2)+timetk::lag_vec(date_key,lag=3)
    ,.show_summary = TRUE
    # ,show_summary=TRUE
  )

```


## transformation functions

-   provides time based mutate over time windows this is like group_by + mutate
mutate_by_time(
.by="week"
,formula
)

- applies log +1 to a column
loglp() 

- standardize and centers
standarize_vec() 

-   normalize a vector
timetk::normalize_vec()


-   lags a columns

lag_vec( shifts a vector)

-   creates many lags against a column

tk_augment_lag(
.lags=1:10
)

date_col %-time% "1 day"

date_col %+time% "6 weeks"

## slidify

-   Creates function that does a rolling application, like mean, or correlation
-   use anonymous function ~ as function definition and then plug in other parameters
-   apply that function to a name
-   now you can that function as you want normally eg fun_name()

fun_name = slidify(
.f= function that you want ~cor(.x,.y,use="pairwise.complete.obs")
,period= window length
,align=where is the gap
,partial=what to do with partial
)

```{r}

roll_corr_30 <- slidify(
  .align = "center"
  ,.f=~cor(.x,.y)
  ,.period=30
)

df_daily %>% 
  mutate(
    roll_30_cor=roll_corr_30(.x=vol,.y=mean_up)
    ,rev_change=diff_vec(rev,lag=1)
    ,rev_change_prop=rev_change/lag_vec(rev,lag=1)
    ,roll_30_cor_rev=roll_corr_30(.x=rev_change_prop,.y=roll_30_cor)     
         ) %>% 
  ggplot(aes(x=date_key,y=roll_30_cor_rev))+
  geom_line()+
  # geom_line(aes(y=rev_change_prop),col="red")+
  scale_x_date(date_breaks = "1 month") 
  plotly::ggplotly()

  

```





slidify_vec()

- similar put to a vector


```{r}
slidify_vec(.x = df_daily$vol,.f=~cor(.x,df_daily$mean_up),period=10,.align = "left")


cor(df_daily$vol,df_daily$mean_up)


prep_df <- df_daily %>% 
  recipe(rev~.,data=.) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_pca(all_numeric_predictors()) %>% 
  step_rm(date_key) %>% 

  prep()
  
juice_df =prep_df%>% 
juice()


tidy_df <- prep_df %>% tidy()

tidy_df

df_daily %>% 
  recipe(rev~.,data=.) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_pca(all_numeric_predictors()) %>% 
  step_rm(date_key) %>% tidy()


df %>% 
  group_by(store_key) %>% 
  summarise(
    rev=sum(rev)
    ,vol=sum(sales_quantity)
    ,returns=sum(return_amount)
    ,discounts=sum(discount_amount)
    ,.groups = "drop"        
    )  %>% 
  mutate(across(2:5,~.x/sum(.x))) %>% 
  recipe(store_key~.,data=.) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_pca(all_numeric_predictors(),num_comp = 3) %>% 
  prep() %>% juice() %>% 
  pivot_longer(-1) %>% 
  mutate(
    value=value
    ,store_key=factor(store_key)
    ) %>% 
  filter(store_key %in% 1:10) %>% 
  ggplot(aes(y=store_key,x=value,fill=store_key))+
  geom_col(show.legend = FALSE)+
  facet_wrap(~name,nrow = 1,scales = "free")
step_pca()


df %>% 
  nest_by(store_key) %>% 
  mutate(
    data=list(data)
    ,n=nrow(data)
    ,rev=sum(data$rev)
    ,vol=sum(data$sales_quantity)
    ,lm_mod=list(lm(rev~sales_quantity,data=data))
  ) %>% 
  mutate(
    rev_prop=rev/sum(rev)
    ,cum_rev_prop=cumsum(rev_prop)
    ,tidy_lm=list(broom::tidy(lm_mod) %>% pull(term))
    ) %>% 
  unnest(tidy_lm)


```


-   creates differences between a log
-   typically used with cumsum for changes in a series
-   lag is how far you want to go back
-   diff =1 is simple lag (period over period)
-   diff=2 is the change in your lagged values, it is the acceleration of difference
diff_vec(
lag
,difference=1 for normal 2 for acceleration
)

```{r}
df_daily %>% 
  mutate(vol_diff=diff_vec(vol,lag=1,diff=1)
         ,vol_acc=diff_vec(vol,lag=1,diff=2)
         ,lag_one=vol-lag(vol,n=1)
         ,lag_of_lag=lag_one-lag(lag_one,1)
           )


343+756
```


- creates forier funcitons, similiar to lag
forier_vec(
period = lag
k=offset
type=cos,sin
)



## Signature
- creates index
tk_index() 

makes time table similiar to create date

tk_make_timeseries() 

- makes a holiday sequence signatures

tk_make_holiday_sequence(
year_start
year_end
calendar
)

tk_get_holiday_signature()



makes future time series
ts_make_future_time_zeries()

future_frame(
lengt_out=periods
)

# Framework

## Extend forcast window

- scale & normalize variables

- create horizon
-   create rolling lag based on horizon
-   create rolling periods

basically make sure our dataframe has everything that it needs for future projects (eg. lags and rolling averages)
- tip create rolling lag of forecast period

- add window

bind_rows(
dataframe,
dataframe %>% 
future_frame(
.data_var
.length_out)
)

- add lags


tk_augment_lags(
variable
.lags
)

- add rolling averages
tk_augment_slidify(
.value=thing that you want rolling average
.f=mean or whatever
.period= c(30,60,90)
.align="center"
.partial=true
)

- add events

anonlomies and external regressors

- save forecast table as seperate table and historical table as seperate


## split and train

-take historical dataset and assign it to object this is your splits

time_series_split(
assess=time horizon
cumulative=TRUE
)

- validate teh splits are right you can use:

tk_time_series_cv_plan() %>% 
plot_time_seris_cv_plan()


## recipe

- can create a base recipe

- to see how these recipes happen use prep() %>% juice() to apply transformats to the dataset

recipe(
formula
data)
%>% 

### creates time series seinguatre
step_timeseries_siganture() #
)

### normalize

- consider to normalize time series signatures 

step_normalize()

### dummy variables
step_dummy(one_hot=TRUE)

### interaction
step_interaction(~ (col)*(col))

### foriere
- base foriere on lags

step_fourier(date,periods=c(),K=2)

### spline
- need to remove lag and date column if you do this
- step_ns(
ends_with(index.num),deg_free=2
)

### lag

-   make sure to remove nas with step_naomit
-   lags added with tk_augment_lags


## workflow

-workflow are model management tools
-   workflows are data dependent due to reciept spec dependecy
-   new data must have the same structure as the unstranformed recipe data
-   worlfow should be created for each combination of model & recipe
-   fitting a wroflow trains the model & recipe on data

- add model and recipe and can fit
- ensure fit is based on same data as reciept (training)

add_model()
add_reciepe()
fit()
## modeltime

- how you compare and forecast models
- modeltime only works with fitted models that have parsnip and workflow
- basically add to to modetime so you acollection of workflows
-   then caolbirate against a dataset (testing)
    
    -   then forecast (against against testing set)
    -   or get accuracy

-add worlflow to model modeltime table
modeltime_table(workflow1,workwlow2)

you can rename model IDs with update_model_description()

- to apply model to different data set (testing),  you use modeltime_calibrate
- ensure you same the calibrated table as a new object
modeltime_calibrate(new_data=testing(split))


-   use modeltime_forecast to apply model to a forecast (pass calibration table\)
-   Need to be mindful that external regressors have column in the forecast table for models
modeltime_forecast(
new_data=testing(splits)
,actual_date=actual_data
,conf_interval=0.0
h= "8 weeks"is horizon can only be used if no external regressors
) %>% 
table_modeltime_accuracy(.interactive=FALSE % GT table) # prints to table

- to get accuracy use modeltime_accuracy() (pass calibration data through to it)
modeltime_acurracy(
metric_set=default_forecast_accuracy_metric_set() for defaul metrics or
metrict_set=yardstick::metric_set(list metrics)
)

- plot model time forecast
plot_modeltime_forecast(
.lengend_max_width=25
)



### refit
-     use to pass model specs to new data source like your future forecasted dataset
-     can use to train data on full dataset
-   pass calibration table to modeltime_refit
calibration_tbl %>% 
modeltime_refit(data=originaltbl)

### residuals

modeltime_residuals() lets you look at the residaual sof the moel


## transfer and revert back

- keep track of parateters that were used to transfrom (log) or normalize data
- make sure you do it in the right order

standarize_inv_vec() to undon standarization

log_interval_inv_vec() to undo log


## Extract elements
-   to pull fit of a workflow use pull_workflow_fit %>% pluck (fit) %>% summary(0)
-   this will tell you how the applied workflow fit looks like


## compare models
-   add wor




## model types

### arima Regression
arima_reg() %>% 
seg_enging("auto_arima")


-   order max tells you how many coefficients you want
-   thee lags are the same as lm(var ~ lag(var,1)+lag(var,2))


```{r}


library(tidyverse)
library(timetk)


df <- fpaR::contoso_fact_sales %>% 
  janitor::clean_names() %>% 
  mutate(
    revenue=unit_price*sales_quantity
   ,date=mdy(date_key) 
    ) 



daily_sales_tbl <- df %>% 
  relocate(last_col()) %>%
  timetk::summarise_by_time(date
                            ,.by = "day"
                            ,n= n()
                            ,revenue=sum(revenue)
                            ,quantity=sum(sales_quantity)
                            ,store_count=length(unique(store_key))
                            ,rev_per_store=revenue/store_count
                            ,quantity_per_store=quantity/store_count
                            )


augment_daily_sales_tbl <- daily_sales_tbl %>% 
  # plot_acf_diagnostics(
  #   .date_var = date
  #   ,.value = rev_per_store
  # )

  mutate(across(-date,timetk::standardize_vec)) %>% 
  tk_augment_lags(
    revenue
    ,.lags = 1:5
                  ) %>% 
  tk_augment_lags(
    quantity
    ,.lags = 1:5
                  ) %>% 
  tk_augment_lags(
    rev_per_store
    ,.lags = 1:5
                  ) %>% 
  timetk::tk_augment_timeseries_signature(date)



augment_daily_sales_tbl %>%
  timetk::tk_augment_holiday_signature(.date_var = date
                                        ,.holiday_pattern = "US_"
                                       ,.locale_set = "US",
                                       .exchange_set="none")




augment_daily_sales_tbl %>% 
  bind_rows(
    future_frame(date_var=date
      ,.length_out = 30

    )
    
  )
bind_rows(
augment_daily_sales_tbl,
augment_daily_sales_tbl %>% 
  future_frame(.date_var = date
               ,.length_out = 30)
)

```



```{r}
#| eval: false

plot_acf <- daily_sales_tbl %>% 
  plot_acf_diagnostics(.date_var,.value) 

plot_anomaly <- daily_sales_tbl %>% 
  plot_anomaly_diagnostics(.date_var,.value)

plot_seasonal <- daily_sales_tbl %>% 
  plot_seasonal_diagnostics(
    .date_var
    ,.value
    )

daily_sales_tbl %>% 
  timetk::tk_augment_timeseries_signature(.date_var) %>% 
  mutate(across(where(is.ordered),as.character)) %>%
  timetk::mutate_by_time(.date_var = .date_var
                         ,.by ="week"
                         ,.type = "round"
                         ,weekly_rev=sum(.value)
                         ,weekly_mean=mean(.value)
                         ,weekly_median=median(.value)
                         ) %>% 
  relocate(last_col()) %>% 
  mutate(across(contains("weekly_")
                ,standardize_vec)) %>% 

  
  mutate(.value=log(.value+1),
         
  ) %>% view()
  lm(.value ~.-.date_var
     ,data=.) %>% 
  summary()
  broom::tidy() %>% 
  filter(p.value<.05)
  timetk::plot_time_series_regression(.date_var,
                                      .value~.)


```


```{r}
library(tidymodels)
split <- daily_sales_tbl %>% 
  timetk::time_series_split(
    date_var =date
    ,assess = 8
    ,cumulative = TRUE
    )

train <- rsample::training(split)
test <- rsample::testing(split)



lm_mod <- parsnip::linear_reg()
ph_mod <- modeltime::arima_boost()


rec_basic <- recipes::recipe(revenue~.,data=train) %>% 
  timetk::step_timeseries_signature(date) %>% 
  timetk::step_fourier(date,period =c(1,3,7),K = 2)


rec_basic 


lm_basic <- workflow() %>% 
  add_model(lm_mod) %>% 
  add_recipe(rec_basic) %>% 
  fit(train)

ph_basic <-  workflow() %>% 
  add_model(ph_mod) %>% 
  add_recipe(rec_basic) %>% 
  fit(train)

modeltime::modeltime_table(
  lm_basic
  ,ph_basic
) %>% 
  modeltime::modeltime_calibrate(new_data = train) %>% 
  modeltime::modeltime_accuracy()



```

