---
title: "time series"
---


framework

1. summarize by date column, clean and diagnostics
2. add dimensions that you want in as input or to frame by learning
3. add correlations such as lag or other dimensions
4. scale and normalize data, including log
5. add time related signatures






## how to create aggreation fucntion
-   Use `summarize_by_time()` to aggregate data by a time dimension 
-   Similar to group_by()+summarize() but with specialized features for time
        
      -   .date_var=date column
      -   .by= "day", "week", "month", "quarter", "year", "seaon", "bimonth", "halfyear"
      -   .type=c("ceiling","round","floor")
      -   .week_start(1=monday, 7=sunday)
      -   col_name=aggregation step (n=n()) or across()
  

## create diagnostics of summarized by time model 

-   Creates summary table that identifies differences between time series and any gaps as well summarize basic state

    -   start_date
    -   end_date
    -   units
    -   timezone
    -   distribution of timer series difference in second

##  Clean data: how to pad time when there gaps in data series
-   use `pad_by_time()` to fill gaps with values in your time series

    -   .date_var=data column
    -   .by="day"
    -   .pad_value=0

## plot times series

-  `plot_time_series()` to  plots time series either in plotly, ggplot, or trelliscope
-   Can add groups by passed in a grouped dataframe
-   Useful for EDA 

    -   .data_var=date column
    -   value = the value you want to plot
    -   .color_bar=group_name # puts groups in color (or you can put this group_by)
    -   .facet_vars=name # also how to facet
    -   .smooth=TRUE (adds in polynominal regression)
    -   .smooth_period="30 days"
    -   .smooth_span= prop of data
    -   .interactive=FALSE # to create ggplot object



## Filter time series 

filter_by_time(
.date_var=data column
start_date=
end_date="end"  or date
)

## split into training and test set

- you assign the split data to an object similiar to tidymodels

 time_series_split(
 date_var=date colomns
 assess= "8 weeks"
 cumulaitve = TRUE  sets the previous as training, false means you need to select training set
)

then you use the tk_time_series_cv_plan()  and plot_time_Series_cv_plan() to see the training and test data 


## create different time series models

linear_reg(
)

prophet_reg(
) %>% 
set_engine("prophet")


how to use?

prophet_reg %>% 
set_engine("prophet") %>% 
fit(predition_Value ~ data_col
,new_data=training(splits))


## model time worfklow
creates table to hold various fitted models
model_time_table(
fitted_model_1
)

- applies fitted models against testing data
- .calabration_data has the results
- savve as calibration data
model_tbl %>% 
modeltime_calibrate(
new_data=testing(splits)

)

to see this in practice
calibration_tbl %>% 
modeltime_forecat(actual_Date=existing_tlb) %>% 
plot_modeltime_forecast()

## how to see forecasting accuracy

calibration_tbl %>% modeltime_accuracy


plot_seasonal_diagnotistic


## recipe

recipe(formula,training(splits)) %>% 
step_timeseries_Signature() %>%
step_normalize(index.num,_year) %>% 
step_dummy(all_ordinal)
prep()
%>% juice (applies to the training step)


## create model element

linear_reg()

## create a workflow object
save as worflow

workflow %>% 
add_model(model_spec) %>% 
add_reciept(reciepe-sepc) %>% 
fit(training(splits)) 

calibration obj
modeltime_table() %>% 
modeltime_Calibration(testing(splits))


calibration_obj %>% modeltime_accuracy

calibration_obj %>% 
modeltime_forecast(
new_Data=testing(splits)
actual_Data=origional_ttbl) %>% 
plot_modeltime_forecast()



)

# overview of functions

## diagnostics functions

### ACF

-   helpful to detect lagged functions and how many lags to use

plot_acf_diagnoistcs(
.lags="1 year", 25:100,100
.ccf_vars=columns:columns   
.show_ccf_vars-only=TRUe # ony cross correlations var
,.facet_ncol=number of column

)


## sesaonality

-   Helpful to see if there are certain seasonality drivers

plot_seaonsal_diagnotistc(
feasture_set=c(hour,weekday)
.geom="violin"
)

### anolomy

-   helpful to detect anomoly detection

plot_anomoly_diagonsitics(
.data_var=data_col
.alpha= .01 #change anomoly sensitivity
.max_anolomoies =.01 # peak anomoly

)

-   Get underlying anomaly data
tk::anomoly_diaogasitc(
.data_var=data_col
.alpha= .01 #change anomoly sensitivity
.max_anolomoies =.01 # peak anomoly
)

## seasons decomposition

-   Useful to understand how anomoliges are being detcted
plot_stl_diagostics(


)


## linear regression plot

-   shortcut function of `lm()` to help identify which features are meaningful or not

plot_time_series_regression(
.date_var=data column
.formula= value~ as.numeric(date_col)+
wday(date_col)+
month(date_col)
show_summary= TRUE, shows summary ouptut
)

## transformation functions

-   provides time based mutate over time windows this is like group_by + mutate
mutate_by_time(
.by="week"
,formula
)

- applies log +1 to a column
loglp() 

- standardize and centers
standarize_vec() 

-   normalize a vector
timetk::normalize_vec()


-   lags a columns

lag_vec( shifts a vector)

-   creates many lags against a column

tk_augment_lag(
.lags=1:10
)

date_col %-time% "1 day"

date_col %+time% "6 weeks"



slidify(
.f= function that you want ~cor(.x,.y,use="pairwise.complete.obs")
,period= window length
,align=where is the gap
,partial=what to do with partial
)

slidify_vec()

- similar put to a vector


-   creates differences between a log
-   typically used with cumsum for changes in a series
diff_vec(
lag
,difference=1 for normal 2 for acceleration
)

- creates forier funcitons, similiar to lag
forier_vec(
period = lag
k=offset
type=cos,sin
)



## Signature
- creates index
tk_index() 

makes time table similiar to create date

tk_make_timeseries() 

- makes a holiday sequence signatures

tk_make_holiday_sequence(
year_start
year_end
calendar
)

tk_get_holiday_signature()



makes future time series
ts_make_future_time_zeries()

future_frame(
lengt_out=periods
)

# Framework

## Extend forcast window

- scale & normalize variables

- create horizon
-   create rolling lag based on horizon
-   create rolling periods

basically make sure our dataframe has everything that it needs for future projects (eg. lags and rolling averages)
- tip create rolling lag of forecast period

- add window

bind_rows(
dataframe,
dataframe %>% 
future_frame(
.data_var
.length_out)
)

- add lags


tk_augment_lags(
variable
.lags
)

- add rolling averages
tk_augment_slidify(
.value=thing that you want rolling average
.f=mean or whatever
.period= c(30,60,90)
.align="center"
.partial=true
)

- add events

anonlomies and external regressors

- save forecast table as seperate table and historical table as seperate


## split and train

-take historical dataset and assign it to object this is your splits

time_series_split(
assess=time horizon
cumulative=TRUE
)

- validate teh splits are right you can use:

tk_time_series_cv_plan() %>% 
plot_time_seris_cv_plan()


## recipe

- can create a base recipe

- to see how these recipes happen use prep() %>% juice() to apply transformats to the dataset

recipe(
formula
data)
%>% 

### creates time series seinguatre
step_timeseries_siganture() #
)

### normalize

- consider to normalize time series signatures 

step_normalize()

### dummy variables
step_dummy(one_hot=TRUE)

### interaction
step_interaction(~ (col)*(col))

### foriere
- base foriere on lags

step_fourier(date,periods=c(),K=2)

### spline
- need to remove lag and date column if you do this
- step_ns(
ends_with(index.num),deg_free=2
)

### lag

-   make sure to remove nas with step_naomit
-   lags added with tk_augment_lags


## workflow

-workflow are model management tools
-   workflows are data dependent due to reciept spec dependecy
-   new data must have the same structure as the unstranformed recipe data
-   worlfow should be created for each combination of model & recipe
-   fitting a wroflow trains the model & recipe on data

- add model and recipe and can fit
- ensure fit is based on same data as reciept (training)

add_model()
add_reciepe()
fit()
## modeltime

- how you compare and forecast models
- modeltime only works with fitted models that have parsnip and workflow
- basically add to to modetime so you acollection of workflows
-   then caolbirate against a dataset (testing)
    
    -   then forecast (against against testing set)
    -   or get accuracy

-add worlflow to model modeltime table
modeltime_table(workflow1,workwlow2)

you can rename model IDs with update_model_description()

- to apply model to different data set (testing),  you use modeltime_calibrate
- ensure you same the calibrated table as a new object
modeltime_calibrate(new_data=testing(split))


-   use modeltime_forecast to apply model to a forecast (pass calibration table\)
-   Need to be mindful that external regressors have column in the forecast table for models
modeltime_forecast(
new_data=testing(splits)
,actual_date=actual_data
,conf_interval=0.0
h= "8 weeks"is horizon can only be used if no external regressors
) %>% 
table_modeltime_accuracy(.interactive=FALSE % GT table) # prints to table

- to get accuracy use modeltime_accuracy() (pass calibration data through to it)
modeltime_acurracy(
metric_set=default_forecast_accuracy_metric_set() for defaul metrics or
metrict_set=yardstick::metric_set(list metrics)
)

- plot model time forecast
plot_modeltime_forecast(
.lengend_max_width=25
)



### refit
-     use to pass model specs to new data source like your future forecasted dataset
-     can use to train data on full dataset
-   pass calibration table to modeltime_refit
calibration_tbl %>% 
modeltime_refit(data=originaltbl)

### residuals

modeltime_residuals() lets you look at the residaual sof the moel


## transfer and revert back

- keep track of parateters that were used to transfrom (log) or normalize data
- make sure you do it in the right order

standarize_inv_vec() to undon standarization

log_interval_inv_vec() to undo log


## Extract elements
-   to pull fit of a workflow use pull_workflow_fit %>% pluck (fit) %>% summary(0)
-   this will tell you how the applied workflow fit looks like


## compare models
-   add wor




## model types

### arima Regression
arima_reg() %>% 
seg_enging("auto_arima")

```{r}
#| eval: false

library(tidyverse)
library(timetk)


df <- fpaR::contoso_fact_sales %>% 
  janitor::clean_names() %>% 
  mutate(
    revenue=unit_price*sales_quantity
   ,date=mdy(date_key) 
    ) 



daily_sales_tbl <- df %>% 
  relocate(last_col()) %>%
  timetk::summarise_by_time(date
                            ,.by = "day"
                            ,n= n()
                            ,revenue=sum(revenue)
                            ,quantity=sum(sales_quantity)
                            ,store_count=length(unique(store_key))
                            ,rev_per_store=revenue/store_count
                            ,quantity_per_store=quantity/store_count
                            )


augment_daily_sales_tbl <- daily_sales_tbl %>% 
  # plot_acf_diagnostics(
  #   .date_var = date
  #   ,.value = rev_per_store
  # )

  mutate(across(-date,timetk::standardize_vec)) %>% 
  tk_augment_lags(
    revenue
    ,.lags = 1:5
                  ) %>% 
  tk_augment_lags(
    quantity
    ,.lags = 1:5
                  ) %>% 
  tk_augment_lags(
    rev_per_store
    ,.lags = 1:5
                  ) %>% 
  timetk::tk_augment_timeseries_signature(date)



augment_daily_sales_tbl %>%
  timetk::tk_augment_holiday_signature(.date_var = date
                                        ,.holiday_pattern = "US_"
                                       ,.locale_set = "US",
                                       .exchange_set="none")




augment_daily_sales_tbl %>% 
  bind_rows(
    future_frame(date_var=date
      ,.length_out = 30

    )
    
  )
bind_rows(
augment_daily_sales_tbl,
augment_daily_sales_tbl %>% 
  future_frame(.date_var = date
               ,.length_out = 30)
)

```



```{r}
#| eval: false

plot_acf <- daily_sales_tbl %>% 
  plot_acf_diagnostics(.date_var,.value) 

plot_anomaly <- daily_sales_tbl %>% 
  plot_anomaly_diagnostics(.date_var,.value)

plot_seasonal <- daily_sales_tbl %>% 
  plot_seasonal_diagnostics(
    .date_var
    ,.value
    )

daily_sales_tbl %>% 
  timetk::tk_augment_timeseries_signature(.date_var) %>% 
  mutate(across(where(is.ordered),as.character)) %>%
  timetk::mutate_by_time(.date_var = .date_var
                         ,.by ="week"
                         ,.type = "round"
                         ,weekly_rev=sum(.value)
                         ,weekly_mean=mean(.value)
                         ,weekly_median=median(.value)
                         ) %>% 
  relocate(last_col()) %>% 
  mutate(across(contains("weekly_")
                ,standardize_vec)) %>% 
  
  mutate(.value=log(.value+1),
         
  ) %>% view()
  lm(.value ~.-.date_var
     ,data=.) %>% 
  summary()
  broom::tidy() %>% 
  filter(p.value<.05)
  timetk::plot_time_series_regression(.date_var,
                                      .value~.)


```


```{r}
#| eval: false

library(tidymodels)
split <- daily_sales_tbl %>% 
  timetk::time_series_split(
    date_var =date
    ,assess = 56
    ,cumulative = TRUE
    )

train <- rsample::training(split)
test <- rsample::testing(split)



lm_mod <- parsnip::linear_reg()
ph_mod <- modeltime::prophet_reg()


rec_basic <- recipes::recipe(revenue~.,data=train) %>% 
  timetk::step_timeseries_signature(date) %>% 
  timetk::step_fourier(date,period =c(1,3,7),K = 2)


rec_basic 


lm_basic <- workflow() %>% 
  add_model(lm_mod) %>% 
  add_recipe(rec_basic) %>% 
  fit(train)

ph_basic <-  workflow() %>% 
  add_model(ph_mod) %>% 
  add_recipe(rec_basic) %>% 
  fit(test)

modeltime::modeltime_table(
  lm_basic
  ,ph_basic
) %>% 
  modeltime::modeltime_calibrate(new_data = train)



```





