---
title: "Graphics with ggplot2"
author: "Alejandro"
date: "11/21/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


setting github
use_git_config(user.name = "haganistic", user.email = "alejandro.hagan@outlook.com")


# Admin how to set up your working space

## Manipulating working directories

getwd()

setwd()

list.files(pattern=$^, recursive=T)


`.` is equal to your working director `getwd()`


## Creating folders & filing guidelines 

file.exists
dir.exists
basename("c:/filedir/file.txt") -> "file.txt
dirname("c:/filedir/file.txt")->

create.dir("./newdir")
file.create("filename.txt")
shell.exec("filename.txt")
filesstrings::move_file("c:home/filename.txt","c:newdir")
file.copy("filename", "destination")
file.rename(directory,"existing file name","new file name")
-if directory is blank will assume getwd()
file.choose()
base::unlink("filename", recursive=T)
-recursve must be T to delete directories

[Link to learn more](http://theautomatic.net/2018/07/11/manipulate-files-r/#:~:text=Files%20can%20be%20deleted%20with%20R%20using%20unlink.,parameter%20recursive%20%3D%20TRUE.%20%23%20delete%20a%20file)


## how to connect to github

while all this may seem to new, quite honestly its because we have never been told good versioning and collaboration practices so we are reliant on saving to a LAN system, amending the name with  versioning control


## push & committ and commenting


# How to manipulate & tidy data

## how to import data

read.csv()

read_csv()

fread()

## how to change column types





 



## how to create data
seq()
runif()
sample(data,# of times, replace,)
c()
data.frame()

## how to create data with basic loops / repitition

## types of data

vector
data.frame
list


## some base R basics that will be helpful to you as you read the forums

## how to import files

- by folder
- by website
- excel spreadsheet
- powerbi model
- 

### how to find files by their type

### how to append files together

### how to automate file importation 


## file column names

### how to change column names

#### statically


#### dynamically

#### best practices when naming columns

### how to change column types

#### statically

#### dynamically

### how to check data structure
-unique

### how to clean data



### how to shape data


### how to subset data


#### filter

dynamic

static

### select

dynamic

static

### grouby


### summarize


### mutate


### fill / blanks/ replace values


### recode


### pivot / unpivot


### merge


### iterate


- apply a single function (single input or multiple inputs) to every/ some/ based on criteria column and get its outputs into a data frame (to be accessed)

- apply a single function with multiple inputs

- apply a functions to nested data frames

### explatory analysis 

- start with variance analysis 



# how to extract information out of a table

for (i in range){

do something
}

basic framework 
1. is to define the vector that you want as an output and its size with `vector("type",vectorsize)`
2. assign that vector names with `names()` so that data has headings
3. create for (i in range)
4. define the task it will do individually to thedataset
4. assign that to the output vector


tips & tricks

- check the task individually against a single element to ensure it works
- you can extract other elements (typically row / column names) using other techniques (names()) and assign to output (doesn't need to be in the loop)
-use `seq_along()`, `ncol()`, `[[]]` to define parameters and extract elements


``` {r loop_example}
output <- vector("double",ncol(mtcars)) #assigns the output vector

cols_names <- names(mtcars) # assigns names of the dataframe to variable
names(output) <- cols_names # assigns the names 


for (i in cols_names) { # sometimes useful to use seq_along() here as well
  
  
  output[i] <- mean(mtcars[[i]]) # the double brackets ensures we only take out one item, this needs to be adjusted if two iems are expected, change the vector to a list
  
}

output

```
 
 ### Categorical Variables
 
 In general you will need to distinguishbetwen your character values as either straight character or categorical with levels
 
 This becomes critical as you look to create categories and relationships in your data
 
 forecats is the gotopackage in particular:
 

1) rename

recode() to change values in column

`recode(col,newvalue=oldvalue)`

2) reorder

fct_relevel

`fct_relevel(col,level1,level2,etc)`

3)group variables into another group

case_when()

-typically used in combination with mutate()
-can reference multiple conditions

`case_when(col1==var1 ~ val1,
            col1==var2 & col3==var3 ~ var 2,
            is.na(col1) ~ "missingvalue",
            TRUE ~ "defaultvalue`
            )`
cut()





