---
title: "r_for_data_science_answers"
author: "Alejandro"
date: "11/27/2021"
output: html_document
---

```{r setup, include=FALSE}

install.packages("pacman")
pacman::p_load(tidyverse,filesstrings)
knitr::opts_chunk$set(echo = TRUE)
```

# 3.2.4

*Run ggplot(data = mpg). What do you see?

we don't see anything because no geoms have been mapped

*How many rows are in mpg? How many columns?

There are `r dim(mpg)[1]` rows and `r dim(mpg)[2]` columns

*Make a scatterplot of hwy vs cyl.


```{r 3.2.4 scatterplot}


mpg %>% ggplot(aes(x=hwy, y=cyl))+geom_point()

```


*What happens if you make a scatterplot of class vs drv? Why is the plot not useful?


plot isn't useful because they are both character type variables, a count of each of proportion would be more helpful

```{r 3.2.4 scatterplot2}


mpg %>% ggplot(aes(x=class, y= drv))+
geom_point()

```


# 3.3.1

*What’s gone wrong with this code? Why are the points not blue?

points are blue because "blue" is mapped as an aes()

*Which variables in mpg are categorical? Which variables are continuous?

``` {r class of mpg }


str(mpg) %>% knitr::kable("pipe")

```


*Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?


``` {r class vs. class}

# continuous vs.continuous  
mpg %>% ggplot(aes(x=.$hwy,y=.$cty, color=.$class, size=.$cyl, shape=.$class))+geom_point()

```



*What happens if you map the same variable to multiple aesthetics?


``` {r multiple mappings}


mpg %>% ggplot(aes(x=cty, y=hwy, col=class,size=class,shape=class))+
  geom_point()



```


*What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)

alters color line thickness

``` {r stroke}


mpg %>% ggplot(aes(x=cty, y=hwy))+
  geom_point(aes(stroke=0.4))

```

*What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.v


``` {r scatter filter}


mpg %>% ggplot(aes(x=cty, y=hwy))+
  geom_point(aes(stroke=0.4))

aes(colour = displ < 5)

```


